# Practice Questions - TOPIC: Vectors2

## Vectors

### QUESTION 1 *(vectors, rep/seq/:/etc)*

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION:
#
# Write an R command that displays all the even numbers between 100 and 1000 (including 100 and 1000) 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
seq(100,1000,by=2)  #<<ANSWER>>
```
:::

---

### QUESTION 2 *(vectors , rep/seq/:/etc)*

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION:
#
# Write an R command that displays the values:   10, 13, 16, 19, 22, etc. There
# should be 500 numbers in all.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
seq(10,by=3,length.out=500)  #ANSWER
```
:::

---

### QUESTION 3. *(vectors , rep/seq/:/etc)*

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION
#
# Write an R command that displays alternating values of positive and
# negatives of 1,2,3 ie.:  1,2,3,-1,-2,-3,1,2,3,-1,-2,-3,1,2,3 ... etc.
# There should be a total of 60 numbers in all.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
rep(c(1:3,-1:-3),length.out=60)   # ANSWER

rep(c(1,2,3,-1,-2,-3),length.out=60)   # ANOTHER ANSWER
```
:::

---

### QUESTION 4. *(vectors , recycling)*

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 4    TOPICS: vectors , recycling  ####
# 
# Write an R command that displays all the powers of 2 from 
# 2 to the power of minus 20 (i.e. -20) until
# 2 to the power of positive 20 (i.e. +20)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
-20:20      # this gives the numbers -20 -19 -18 ... -2 -1 0 1 2 ... 18 19 20

2^(-20:20)  # ANSWER

2^-20:20    # NOTE: This is WRONG because order of operations makes this work like (2^-20):20
```
:::

---

### QUESTION 5 *(vectors , indexingWithLogicalValues)*

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 
#
# Write an R command that displays every entry in the vector x whose value is 10 or more.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(3,10,11,2,-5,100,21)  # setup some example data
x
x >= 10       # This gives TRUE/FALSE values - there will be one 
              # TRUE or FALSE for each value in x. Given the data above,
              # this will be: c(FALSE,TRUE,TRUE,FALSE,FALSE,TRUE,TRUE)

x[x>=10]      # ANSWER - The first "x" outside the [brackets] is the vector
              #          whose values are going to be displayed. The 
              #          "x>=10" inside the [brackets] is used to generate
              #          a logical vector (i.e. TRUE/FALSE values) to indicate
              #          which values will be displayed. Only those values
              #          from x whose positions correspond to a TRUE value
              #          in the logical vector, will be displayed. Given the
              #          sample data above, the 2nd, 3rd, 5th and 7th values
              #          will be displayed, i.e. 10,11,100,21
```
:::

---

### QUESTION 6    TOPICS: vectors , indexingWithLogicalValues ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 6    TOPICS: vectors , indexingWithLogicalValues ####
#
# x and y are two numeric vectors whose lengths are equal.
# Show every value from the vector, x, that is less than the corresponding value in the vector y
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(3,10,11,2,-5,100,21)  # setup some example data 
y <- c(10,1,11,15,5,9,999)   # setup some example data
length(x)
length(y)
x < y         # This gives TRUE/FALSE values - there will be one 
              # TRUE or FALSE for each value in x. Given the data above,
              # this will be: c(FALSE,TRUE,TRUE,FALSE,FALSE,TRUE,TRUE)

x[x<y]     # ANSWER
y[x<y]     # FYI - this wasn't asked but these are the corresponding
           # values from y that are greater than those in x
```
:::

---

### QUESTION 7    TOPICS: vectors , round , runif , RHelpPages  ####


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 7    TOPICS: vectors , round , runif , RHelpPages  ####
#
# SOME BACKGROUND - runif function
#
#     The runif function returns a numeric vector that contains "random numbers
#     with a uniform distribution". The R help page for runif shows that the
#     arguments and default values for those arguments are:
#         runif(n, min=0, max=1)     
#     where
#         n is the number of numbers to generate
#         min is the smallest value that a number could have
#         max is the largest value that a number could have. 
#
#     EXAMPLES: 
#  
#     > # return a vector of 3 random numbers whose values fall between 0 and 10
#     > runif(3,0,10)   
#     [1] 3.785233 9.672244 2.857430
# 
#     > # return a vector of 10 random numbers whose values fall between -5 and +5
#     > runif(10,-5,5)
#     [1] -1.84032071  4.28685602  2.59656498 -1.09880205 -3.90202237 -1.89618831  3.37193245
#     [8]  0.08757248 -1.10848900 -1.90661883
# 
#     > # return a vector of 5 random numbers. The min and max are not specified so 
#     > # use the default values of min (ie. 0) and max (i.e. 1)
#     > runif(5)
#     [1] 0.4098594 0.8771700 0.5610475 0.8928674 0.4863841 
#
#   WHAT YOU HAVE TO DO:
#
#     Generate 10 random numbers that are between 50 and 51. The numbers should have
#     2 values after the decimal point. (HINT: use the   runif   and the   round
#     functions). Note that every time you run this command the actual numbers will
#     be different.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
runif(10,50,51)                  # generate 10 random numbers between 50 and 51

round(runif(10,50,51),digits=2)  # ANSWER 
```
:::

---

### QUESTION 8    TOPICS: vectors , c ####


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 8    TOPICS: vectors , c ####
#
# Insert the number 5 in the beginning of the vector named x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(7,8,2,1)  # start with some data for x (this could be ANY data)
x <- c(5,x)      # ANSWER -  x is now c(5,7,8,2,1)
x             

x[1] <- 5    # WRONG: This will change the first position to 5
# given the example data above the 7 will be overwritten with the 5
# and x will be c(5,8,2,1)
```
:::

---

### QUESTION 9   TOPICS: vectors , c , length  ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 9   TOPICS: vectors , c , length  ####
#
# Insert the number 6 at the end of the vector named x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# ONE WAY
x <- c(7,8,2,1)    # setup some example data
x <- c(x,6)    # ANSWER
x


# ANOTHER WAY
x <- c(7,8,2,1)    # setup some example data
x[length(x)+1] = 6   # ANOTHER ANSWER
x
```
:::

---

### QUESTION 10   TOPICS: vectors , c , indexingWithNumbers

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 10   TOPICS: vectors , c , indexingWithNumbers
#
# x is a vector that contains 11 or more values.
# Insert the number 7 between the 10th and 11th values in the vector named x.
#
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- seq(10,100,by=10)    # setup some example data - x is now:
x <- seq(10,150,by=10)    # setup some example data - x is now:
#   c(10,20,30,40,50,60,70,80,90,100,110,120,130,140,150)
x
x <- c(x[1:10],7,x[11:length(x)])  # ANSWER 
x
#   x is now :
#   c(10,20,30,40,50,60,70,80,90,100,7,110,120,130,140,150)
#
#   The "length(x)" is necessary so that this answer
#   will work no matter how many values the vector, x, contains.
#   the data above was just and example, but the command should 
#   now matter what values x contains. 

x
```
:::

---

### QUESTION 11   TOPICS: vectors , c , indexingWithNumbers ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 11   TOPICS: vectors , c , indexingWithNumbers ####
#
# The vector x contains 10 values. Swap the first 5 and 
#     the last 5 values in x. For example, if x contains the 
#     values 1,2,3,4,5,6,7,8,9,10  then after your command runs, 
#     x should contain the values 6,7,8,9,10,1,2,3,4,5
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- seq(10,100,by=10) # setup example data x contains c(10,20,30,40,50,60,70,80,90,100)

x
x <- c(x[6:10],x[1:5]) # ANSWER -  x now contains c(60,70,80,90,100,10,20,30,40,50)
x
```
:::

---

### QUESTION 12   TOPICS: namedVectors ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 12   TOPICS: namedVectors ####
#
# Grades is a named vector that contains test scores.
#     The names of the entries are the names of the students.
#     Exchange the value of Tom's grade with Joe's grade where
#     "tom" and "joe" are the names of the values that contain
#     the grades for Tom and Joe.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
grades <- c(90,80,70,100,85)                        # setup some example data: set the values in the vector
names(grades) <- c("sue","tom","bill","ann","joe")  # setup some example data: set the names of the values
names(grades)                                       # this displays just the names: "sue","tom",etc.
grades                                              # this displays the values:  90,80,70,100,85
grades[c("tom","joe")]                              # this displays tom's grade and joe's grade: 80  85
grades[c("joe","tom")]                              # this displays joe's grade and tom's grade: 85  80

grades[c("tom","joe")] <- grades[c("joe","tom")]  # ANSWER: this puts the value of 
# Joe's grade into the position for Tom's grade and
# Tom's grade into the position for Joe's grade.
grades                                              # show the new vector
```
:::

---

### QUESTION 13   TOPICS: namedVectors , indexingWithLogicalValues ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 13   TOPICS: namedVectors , indexingWithLogicalValues ####
#
# Grades is a named vector that contains test scores. The names 
# of the entries are the names of the students. Change susan's name to sue
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
grades <- c(90,80,70,100,85)
names(grades) <- c("susan","tom","bill","ann","joe")
grades
names(grades)=="susan"   # Shows TRUE/FALSE values
names(grades) [ names(grades)=="susan" ]  # identify item in the names vector that contains susan

# The following line is the ONLY line that is required for the answer.
# The code puts the value "sue" into the position of the vector, names(grades)
# that currently contains the value "susan"

names(grades) [ names(grades)=="susan" ] <- "sue"       # ANSWER 

grades
```
:::

---

### QUESTION 14   TOPICS: vectors , indexingWithNumbers , vectorAssignment ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 14   TOPICS: vectors , indexingWithNumbers , vectorAssignment ####
#
# Exchange the 4th and 7th value in the vector nums.
#
#      a.  Do this using 3 different commands. Hint, you may create new variables. 
#          (Hint, you may create new variables. For example, I'll start you off:
#              fourth <- x[4]
#              seventh <- x[7]
#              # fill in the rest of the R code to answer the question.
#
#      b. Do this using only a single R command
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
#14a
nums <- seq(10,100,by=10)
x <- nums[7]        # ANSWER (MULTIPLE LINES)
nums[7] <- nums[4]  # ANSWER (MULTIPLE LINES)
nums[4] <- x        # ANSWER (MULTIPLE LINES)
nums

#14b
nums <- seq(10,100,by=10)
nums[c(4,7)] <- nums[c(7,4)]   # ANSWER
nums
```
:::

---

### QUESTION 15   TOPICS: vectors , indexingWithNumbers , length

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 15   TOPICS: vectors , indexingWithNumbers , length
#
# Display the last value from the vector x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(10,20,30,40,50)
x[length(x)]    # ANSWER
```
:::

---

### QUESTION 16   TOPICS: vectors , indexingWithNumbers , length

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 16   TOPICS: vectors , indexingWithNumbers , length
#
# Display the 2nd to last value in x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(10,20,30,40,50)
x[length(x)-1]          # ANSWER
```
:::

---

### QUESTION 17   TOPICS: vectors , indexingWithNumbers , mean , orderOfOperations ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 17   TOPICS: vectors , indexingWithNumbers , mean , orderOfOperations ####
#
# Display the average of the last 5 values in x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(1,5,2,3,1,20,4,10,6,50)
mean(x[(length(x)-4):length(x)])   # ANSWER

# Note, in the answer above, notice the parentheses around (length(x)-4).
# These parentheses are REQUIRED due to order of operations. Without the parentheses
# the command, length(x)-4:length(x) would actually be interpreted by R as 
# length(x)-(4:length(x)) which would give the WRONG result. For more information, 
# see the "recipe" about "order of operations in R" in the "R Cookbook" published by O'Reilly
```
:::

---

### QUESTION 18a   TOPICS: vectors , vectorArithmetic 

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 18a
# TOPICS: vectors, vectorArithmetic, seqColon, indexingWithLogicalValues, length, sciNotation, Inf
#
# Inf is used for "Infinity"
# 
# R can only work with numbers up to a certain size. Any value that is more
# than the maximum is returned as Inf (which stands for Infinity).
#
# Example:
#
# SCIENTIFIC NOTATION
#
# By default R displays very large and very small numbers using a mathematical
# notation called "scientific notation". "Scientific notation" is just a
# fancy way of writing very large and very small numbers using very little
# typing.
#
# To understand "scientific notation", let's start with a simple example.
# 1e+3 is the scientific notation way to write the number 1000.
# 1e+3 can be understood as 1*10^3 (as explained in the next paragraph)
#
# People who are not familiar with scientific notation might read
# "1e+3" as "one e plus 3". However, that is NOT the right way to read it.
# Instead, whenever you see "e" in scientific notation, read it as though
# instead of "e" it said "... times 10 to the power of ...", so "1e+3" should
# be read as "one times 10 to the power of positive 3"
# i.e., "1e+3" is the same as 1*10^3. 
# ( By the way, the letter "e" is used because, "e" is short for "exponent". )
#
# In a similar way, 2.3e+12 is the scientific notation way of writing the
# number 2300000000000 which is the same as "2.3 times 10 to the power of 12"
# i.e. 2.3*10^12
#
# To see that R "understands" scientific notation, just type 1e+3 into R.
# Because 1e+3 is the same as 1000, R just displays 1000. For example:
#
#   > 1e+3      # this is scientific notation for 1000.
#   [1] 1000
#
# By default if a number is small enough then R displays it in it's "regular"
# form (i.e. NOT in scientific notation). However, if a number is very large or
# extremely small, then R displays the number in scientific notation:
#
#   > # "regular numbers" get displayed normally
#   > 1000      
#   [1] 1000
#
#   > # very large numbers get displayed in scientific notation (with a positive exponent)
#   > 1000000000000  
#   [1] 1e+12
#
#   > # extremely small numbers get displayed in scientific notation (with a negative exponent)
#   > 0.0000000000001   
#   [1] 1e-13
#
# -----------------------------------------------------------------------------
# Write code that answers the following question:
#
#    What is the highest power of 4 that is less than one million?
#
# Your code should produce the answer 9. 
#
# The answer is 9 since 4^9=262,144 which is less than 1 million
# but 4^10=1,048,576 is more than one million.
# 
# Write R code to calculate this answer (i.e. 9) automatically.
# Your code should work for any base that is between 2 and 1000 (because 1000 is the
# square root of one million), not only for a base of 4. For example, if instead of
# the number 4, your code used the variable "n", your code should print the
# highest power of "n" that is less than one million. You may write the answer
# using several lines of R code. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
n <- 4               # ANSWER (MULTIPLE LINES)
n2 <- n^(1:1000)     # ANSWER (MULTIPLE LINES)
n2
n2 < 1000000
n2 <- n2[n2<1000000] # ANSWER (MULTIPLE LINES)
n2
length(n2)           # ANSWER (MULTIPLE LINES) (this will display the number 9)
```
:::

---

### QUESTION 18b    TOPICS: vectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 18b
# TOPICS: vectors, vectorArithmetic, seqColon, indexingWithLogicalValues, length
#
# Same as previous question, but this time n contains a "base" as before and
# x contains the maximum value (In the previous question this was one million.
# In this question it can be any number that is contained in x.) The question
# now is what is the maximum whole-number power that n can be raised to whose
# result is less than x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- 200
n <- 3               # ANSWER (MULTIPLE LINES)
n2 <- n^(1:sqrt(x))  # ANSWER (MULTIPLE LINES)
n2 <- n2[n2<x]       # ANSWER (MULTIPLE LINES)
length(n2)           # ANSWER (MULTIPLE LINES)
```
:::

---

### QUESTION 19    TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 19
# TOPICS: vectors, seqColon, indexingWithNumbers, vectorArithmetic
#
# nums is a vector that contains 100 numbers. 
#
# Use nums to create a vector that contains 50 numbers, as described below. 
# The first value should be the sum of the first two numbers from nums. 
# The 2nd value should be the sum of the 3rd and 4th values from nums. 
# For example, if nums contains 1,2,3,4,5,6 ... 99,100, then your answer 
# should contain 3,7,11 ... 199. 
#
# For example, if nums contains
#     3,6,2,8,1,-2, ... ninety-one-other-numbers ... 8,5,-5,2
#
# then your answer should contain
#     9,10,-1, ... forty-five-other-numbers ... 13,-3. 
#
# You may write the answer using several lines of R code.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
nums <- 1:100
firstThirdFifthEtc <- nums[seq(1,100,by=2)]   # ANSWER (MULTIPLE LINES)
secondFourthSixthEtc <- nums[seq(2,100,by=2)]     # ANSWER (MULTIPLE LINES)
answer <- firstThirdFifthEtc + secondFourthSixthEtc  # ANSWER (MULTIPLE LINES)
length(answer)
firstThirdFifthEtc
secondFourthSixthEtc
answer
```
:::

---

### QUESTION 20    TOPICS: vectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 20
# TOPICS: vectors, ==and!=, indexingWithLogicalValues, seqColon
#
# x and y are vectors that each contain 5 numbers. 
#
# PART A:   Show the values from x whose corresponding values in y are the same.
# PART B:   Show the position numbers that contain the same values in x and in y.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
#20a
x <- c(30,50,-20,10,40) # setup some example data
y <- c(99,50,-20,88,40) # setup some example data
x == y                  # This generates a vector that contains five TRUE/FALSE values

x[x==y]                 # ANSWER - "x==y" generates a logical vector (i.e. TRUE/FALSE values).
#          There will be one TRUE or FALSE value for each pair of values
#          in x and y. The "x" outside of the [brackets] indicates the 
#          vector whose values will be displayed. The values from x
#          that occupy the same positions as the TRUE values from "x==y"
#          will be displayed. The values from x that occupy the same positions
#          as the FALSE values from "x==y" will NOT be displayed.


#20b
x <- c(30,50,-20,10,40) # setup some example data
y <- c(99,50,-20,88,40) # setup some example data

(1:length(x))[x==y]     # ANSWER - NOTE: "x==y" generates a logical vector (i.e. TRUE/FALSE values).

#          There will be one TRUE or FALSE value for each pair of values
#          in x and y.
# 
#          The code: "(1:length(x))" that appears outside the [brackets]
#          generates a vector whose values are:
# 
#                   1  2  3  ... etc (up to the length of x).                        # 
#          Given the data above, x contains 5 values. Therefore 
#          the value of (1:length(x))will be:    1  2  3  4  5
# 
#          The code "[x==y]" is used to specify which of the values from the vector
#          should be displayed. The values from "(1:length(x))"
#          that occupy the same positions as the TRUE values from "x==y"
#          will be displayed. The values from x that occupy the same positions
#          as the FALSE values from "x==y" will NOT be displayed.
```
:::
  
---

### QUESTION 21 TOPICS: vectors, sumMean, indexingWithLogicalValues

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 21
# TOPICS: vectors, sumMean, indexingWithLogicalValues
#
# Show all values from the vector x that are greater than the average of the values in x. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(30,50,-20,10,40)
mean(x)
x[x > mean(x)]    # ANSWER
```
:::

---

### QUESTION 22   TOPICS: vectors, sumMean, roundTruncCeilingFloor


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 22
# TOPICS: vectors, sumMean, roundTruncCeilingFloor
#
# The sd() function returns the standard deviation of a vector.
#    Show all of the numbers in the vector x that are at least 1.5 standard deviations
#    away from the mean of x. You can show the numbers in any order you like. They do not
#    have to be in the same order that they appear in the vector x.
#
# HINT:  Standard Deviation is a concept that is covered in statistics. If you didn't take
# stats yet, I can understand why you're confused.  Bottom line, is that sd(c(3,10,76,32,23,31))
# will return a SINGLE number that is called the "standard deviation" of the values
# 3,10,76,32,23,31. sd(c(3,10,76,32,23,31) is 25.67. The mean of those same values
# is 29.16.  Therefore, numbers that are at least 1.5 standard deviations away from the mean are 
#    greater than 29.16 + 1.5 * 25.67    = 67.67
#    or less than 29.16 - 1.5 * 25.67    =  -9.34
# Therefore, if x contained the numbers shown, then you should only display the number 76 as
# that is the only number that is outside of the specified range.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- floor(runif(50,1,100))
x
sd(x)
mean(x)
mean(x) - 1.5 * sd(x)
mean(x) + 1.5 * sd(x)
x[x <= mean(x) - 1.5 * sd(x)]
x[x >= mean(x) + 1.5 * sd(x)]
c ( x[x <= mean(x) - 1.5 * sd(x)]  ,  x[x >= mean(x) + 1.5 * sd(x)]  )    # ANSWER
```
:::

---

### QUESTION 23   TOPICS: vectors, parallelVectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 23
# TOPICS: vectors, parallelVectors
#
# The following variables are parallel vectors that contain info
# about 4 different rectangles.
#
#   colors  <- c("red","blue","green","yellow)
#   lengths <- c(4,5,6,7)
#   widths  <- c(5,6,2,4)
#            
# In other words:
#   the 1st rectangle is "red",    it's length is 4 and it's width is 5
#
#   the 2nd rectangle is "blue",   it's length is 5 and it's width is 6
#
#   the 3rd rectangle is "green",  it's length is 6 and it's width is 2
#
#   the 4th rectangle is "yellow", it's length is 7 and it's width is 4
#
# QUESTIONS:
#
#    The actual numbers and colors shown above are just an example.
#    Your answers must work even if the vectors contain different 
#    numbers and colors and even if the vectors contains fewer or more values.
#
#    a. Show the AREAS of all the rectangles.
#
#            Given the example data above, 
#            your code should evaluate to: 20  30  12 28
#
#    b. Show the COLORS of the rectangles whose areas are at least 25.
# 
#            Given the example data above, 
#            your code should evaluate to: "blue"  "yellow"
#
#    c. Show the COLORS of the rectangles 
#       whose length is at least 5
#       or whose width is at least 5.
#
#           Given the example data above,
#           your code should evaluate to: "red"  blue"  "green" "yellow"
#
#    d. Show the COLORS of the rectangles 
#       whose length and width are both at least 5.
#
#           Given the example data above,
#           your code should evaluate to: "blue"
# 
#    e. Show the COLORS of the rectangles whose area is at least 25.
#
#           Given the example data above,
#           your code should evaluate to: "blue"  "yellow"
# 
#    f. Show the LENGTHS of the blue rectangles whose area is at least 25
#
#           Given the example data above,
#           your code should evaluate to: 5
# 
#    g. Show the LENGTHS of the rectangles whose area is at least 25 
#       and whose color is any one of blue, green or yellow.
#
#           Given the example data above,
#           your code should evaluate to: 5 7
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# 23a.

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

lengths * widths                          # ANSWER



# 23b. 

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

colors[(lengths * widths) >= 25]           # ANSWER




# 23c.

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

colors[lengths>=5 | widths>=5]           # ANSWER


# 23d.

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

colors[lengths>=5 & widths>=5]           # ANSWER


#23e.
# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

colors[lengths*widths>=25]           # ANSWER



# 23f.

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

lengths[colors=="blue" & lengths*widths>=25]           # ANSWER



# 23g. Show the lengths of the rectangles whose area is at least 25 and whose color
#       is any one of blue, green or yellow.

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

# ONE ANSWER: the parentheses are necessary due to the order of operations of & before |
lengths[lengths*widths>=25 & (colors=="blue" | colors=="green" | colors=="yellow") ]    # ANSWER

# ANOTHER ANSWER: this uses the %in% operator
lengths[lengths*widths>=25 & colors %in% c("blue","green","yellow") ]    # ANOTHER ANSWER
```
:::

---

### QUESTION 24   TOPICS: vectors, parallelVectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24
# TOPICS: vectors, parallelVectors
#
# The vector "roomDimensions" is used to store the lengths, widths and
# heights of several rooms in the following format:
# lengthRoom1,widthRoom1,heightRoom1,lengthRoom2,widthRoom2,heightRoom2, etc. 
#
# FOR EXAMPLE, the following code :
#
#         > roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)
#
# represents information about four different rooms with the following dimensions. 
#
#          first room:  length is 10, width is 4, height is 10
#          second room: length is 11, width is 5  height is 9
#          third room: length is 12, width is 5, height is 10.5
#          fourth room: length is 8,  width is 5, height is 9.5
#
# Different numbers may be assigned to roomDimensions, but the structure is as 
# described above. 
#
#    ANSWER PARTS a,b,c,d,e SHOWN BELOW
#      The answer to (a) should display:   4
#      The answer to (b) should display:   11  5  9
#      The answer to (c) should display:   40  55  60  40
#      The answer to (d) should display:   2  3
#      The answer to (e) should display:   9  10
# 
#      NOTE that your code for (a),(b),(c),(d) and (e)  should work no
#      matter what values are placed in the variable, roomDimensions
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

---

#### QUESTION 24a (see the intro to question 24 above)   # TOPICS: vectors, length


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24a (see the intro to question 24 above)
# TOPICS: vectors, length
#
# Write R code to display how many "rooms" are represented in the vector,
# i.e. 1/3 of the number of values in the vector.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# set the roomDimensions vector to some sample data as described in the question
#
#   The first  room is 10ft long, 4ft wide, 10ft tall (ie. the first three numbers in the vector)
#   The second room is 11ft long, 5ft wide, 12ft tall (ie. the next three numbers in the vector)
#   The third  room is 10ft long, 4ft wide, 10ft tall (ie. the next three numbers in the vector)
#   The fourth room is 11ft long, 5ft wide, 12ft tall (ie. the next three numbers in the vector)
roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)

# The number of rooms is 1/3 the number of numbers (since each room is described with 3  numbers)
length(roomDimensions) / 3   # ANSWER
```
:::

---

#### QUESTION 24b (see the intro to question 24 above)   TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24b (see the intro to question 24 above)
# TOPICS: vectors
#
#    Write R code to display the length, width and height of the nth room 
#    where n is a variable that contains a number. For example, if "n" contains 
#    the number 4 then your code should display the 3 numbers that represent the 
#    length, width and height of the 4th room.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# setup some sample data (see the question above)
roomDimensions <- c(10,  4,  10,  11,  5,  9,  12,  5,  10.5,  8,  5,  9.5)

n <- 4    # This is the number of the room we want (i.e. 1 for 1st room, 2 for 2nd room, etc)

# Display the dimensions for just the nth room
#
# To explain the following answer, it helps to think of a particular example.
# Suppose we were trying to get the dimensions for the 4th room ...
# ... keep in mind, that the first number for the 4th room appears right after the last number for the 3rd room.
# ... since there are 3 numbers for each room there are 9 numbers before the first number for the 4th room.
# ... i.e. there are 3*3 = 9 numbers for the first 3 rooms so the 4th room starts at position 10.
# ... to generalize, the first number for the nth room starts at (n-1)*3+1. In the example I just mentioned
# ... n is 4, so (n-1) is 3, so (n-1)*3 is 9 so the first number for the 4th room starts at position
# ... (n-1)*3+1 which is position 10.


roomDimensions[c(10,11,12)]

n=4
roomDimensions[c(10,11,n*3)]


roomDimensions[ c( (n-1)*3+1 , (n-1)*3+2 , (n-1)*3+3 ) ] # ANSWER


# Another way to think of this problem is that the last number for the 4th room is
# at position 4*3 i.e. the 12th number. Since each room as 3 numbers, the first number
# for the 4th room is at position 4*3-1 and the second number for the 4th room is at position 4*3-2.
# Therefore the 3 numbers for the 4th room are at positions 4*3-2 (i.e. 10) , 4*3-1 (i.e. 11) and 4*3 (i.e. 12)
# To generalize, the 3 numbers for the nth room are at positions n*3-2 , n*3-1 and n*3 

roomDimensions[ c(n*3-2, n*3-1, n*3) ]                   # ANOTHER ANSWER
```
:::

---

#### QUESTION 24c (see the intro to question 24 above)  # TOPICS: vectors, parallelVectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24c (see the intro to question 24 above)
# TOPICS: vectors, parallelVectors
#
# Write R code to display the number of square feet of floor space in each room.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# setup some sample data (see the question above)
roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)        

# Get a vector with just the lengths of each room (i.e. 10,11,12,8)
lengths <- roomDimensions[ seq(1,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES)   10 11 12 8
lengths

# Get a vector with just the widths of each room (i.e. 4,5,5,5)
widths <- roomDimensions[ seq(2,length(roomDimensions),by=3)  ] # ANSWER (MULTIPLE LINES) 4 5 5 5
widths

# multiply the lengths and widths to get the square feet of floor space
lengths * widths                                                # ANSWER (MULTIPLE LINES) 40 55 60 40
```
:::

---

#### QUESTION 24d (see the intro to question 24 above) TOPICS: vectors, parallelVectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24d (see the intro to question 24 above)
# TOPICS: vectors, parallelVectors
#
# Write R code to display the room numbers of those rooms that have 
# at least 50 square feet of floor space.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# setup some sample data (see the question)
roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)        

# extract just the lengths into a new vector (10 11 9 10.5)
lengths <- roomDimensions[ seq(1,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES) 10 11 9 10.5
lengths

# extract just the widths into a new vector (4 5 5 5)
widths <- roomDimensions[ seq(2,length(roomDimensions),by=3) ]  # ANSWER (MULTIPLE LINES) 4 5 5 5
widths

# create a vector with the room numbers, i.e. 1 2 3 4
roomNumbers <- 1:(length(roomDimensions)/3)                     # ANSWER (MULTIPLE LINES) 1 2 3 4
roomNumbers

# display the room numbers for those rooms whose length*width is at least 50
roomNumbers[lengths * widths >= 50]                             # ANSWER (MULTIPLE LINES)  2 3
```
:::

---

#### QUESTION 24e (see the intro to question 24 above)  TOPICS: vectors, parallelVectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24e (see the intro to question 24 above)
# TOPICS: vectors, parallelVectors
#
# Write R code to display the heights of those rooms that have 
# at least 50 square feet of floor space.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
######################.
# ONE POSSIBLE ANSWER
######################.
# setup some sample data (see the question)
roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)        

# extract just the lengths into a new vector (10 11 9 10.5)
lengths <- roomDimensions[ seq(1,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES) 10 11 9 10.5
lengths

# extract just the widths into a new vector (4 5 5 5)
widths <- roomDimensions[ seq(2,length(roomDimensions),by=3) ]  # ANSWER (MULTIPLE LINES) 4 5 5 5
widths

# extract just the heights into a new vector (10 9 10.5 9.5)
heights <- roomDimensions[ seq(3,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES)
heights

# get the heights for just the rooms whose area is at least 50 (final answer)
heights[lengths * widths >= 50]                                # FINAL ANSWER (MULTIPLE LINES)  9 10



###########################.
# ANOTHER POSSIBLE ANSWER
###########################.

# setup some sample data (see the question)
roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)        

# extract just the lengths into a new vector (10 11 9 10.5)
lengths <- roomDimensions[ seq(1,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES) 10 11 9 10.5
lengths

# extract just the widths into a new vector (4 5 5 5)
widths <- roomDimensions[ seq(2,length(roomDimensions),by=3) ]  # ANSWER (MULTIPLE LINES) 4 5 5 5
widths

# extract just the heights into a new vector (10 9 10.5 9.5)
heights <- roomDimensions[ seq(3,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES)
heights

# create a vector with the room numbers, i.e. 1 2 3 4
roomNumbers <- 1:(length(roomDimensions)/3)                     # ANSWER (MULTIPLE LINES) 1 2 3 4 
roomNumbers

# find the room numbers for the rooms that have at least 50sq feet of floor space 
desiredRooms <- roomNumbers[lengths * widths >= 50]             # ANSWER (MULTIPLE LINES)  2 3 
desiredRooms

# get the heights for just those rooms (final answer)
heights[desiredRooms]                                           # FINAL ANSWER (MULTIPLE LINES)  9 10
```
:::

---

### QUESTION 25  # TOPICS: vectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 25
# TOPICS: vectors
#
# Display the sum of the squares of the numbers in the vector nums.
# For example, if nums contains the values 4 2 and 5 then your command should
# display the number 45, (i.e. 16 + 4 + 25)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
nums <- c(4,2,5)
sum(nums^2)        # ANSWER
```
:::

---

### QUESTION 26 TOPICS: vectors, indexingWithLogicalValues, vectorAssignment


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 26
# TOPICS: vectors, indexingWithLogicalValues, vectorAssignment
#
# Grades is a vector that contains grades on a test. 
# Modify the vector, grades, by adding 5 points to all grades that are below 60.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
grades <- c(53,70,49,60,90)
grades[grades < 60] <- grades[grades < 60] + 5   # ANSWER
grades
```
:::

---

### QUESTION 27 TOPICS: vectors, indexingWithNumbers

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 27
# TOPICS: vectors, indexingWithNumbers
#
#  Reverse the contents of the vector x. DO NOT USE THE rev FUNCTION. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- seq(10,50,by=10)
x
x <- x[length(x):1]     # ANSWER
x
```
:::

---

### QUESTION 28 TOPICS: vectors, mean, length, indexingWithLogicalValues

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 28
# TOPICS: vectors, mean, length, indexingWithLogicalValues
#
# grades is a vector that contains grades on a test. Display the number of students 
# whose grade is equal to the average grade. For example, if grades contains
# c(80,85,85,85,90) then the answer should be 3 since the average is 85 and 3 students
# received exactly and 85. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
grades <- c(80,85,85,85,90)
mean(grades)
length( grades[grades == mean(grades)] )  # ANSWER
```
:::

---

### QUESTION 29 TOPICS: vectors, seqOrColon, minMax

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 29
# TOPICS: vectors, seqOrColon, minMax
#
# store in the vector nums a sequence of values from the smallest value in the
# vector x to the largest value in x. For example if x would contain c(7,8,7,3,5),
# your code should set the value of nums to the sequence 3 4 5 6 7 8
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(7,10,7,3,5)
min(x)
max(x)
nums <- min(x):max(x)   # ANSWER
nums
```
:::

---

### QUESTION 30 TOPICS: vectors, minMax, c


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 30
# TOPICS: vectors, minMax, c
#
# Write a single command that shows the lowest and the highest values from the
# vector x. For example if x contains c(20,-5,-2,40,25,-2,18), then your answer
# should display -5 40
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(20,-5,-2,40,25,-2,18)    # start with some data for x
c(min(x),max(x))                # ANSWER
```
:::

---

### QUESTION 31 TOPICS: vectors, roundTruncCeilingFloor, sqrt, ==and!= , %/%and%%


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 31
# TOPICS: vectors, roundTruncCeilingFloor, sqrt, ==and!= , %/%and%%
#
# Determine if the number stored in n is a "perfect square". For example 25 is
# a "perfect square" because 5*5 is 25. Similarly, 49 is a perfect square because
# 7*7 is 49. However, 50 is NOT a perfect square since its square root is not an integer.
# Your code should display the value TRUE is n is a perfect square and FALSE if n is
# not a perfect square. Hint: the sqrt(x) function returns the square root of x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
n <- 25                   # let's try with a number that IS a perfect square
trunc(sqrt(n)) == sqrt(n) # ANSWER    (TRUE)
floor(sqrt(n)) == sqrt(n) # ANOTHER POSSIBLE ANSWER   (TRUE)
n%%sqrt(n) == 0           # YET ANOTHER POSSIBLE ANSWER   (TRUE)


# ------- CHECK -------
n <- 30                   # now, just to make sure, let's try with a number that is NOT a perfect square
sqrt(n)
trunc(sqrt(n))
trunc(sqrt(n)) == sqrt(n) # ANSWER   (FALSE)
floor(sqrt(n)) == sqrt(n) # ALTERNATE ANSWER   (FALSE)
n%%sqrt(n) == 0           # YET ANOTHER POSSIBLE ANSWER   (FALSE)
```
:::

---

### QUESTION 32 TOPICS: vectors, roundTruncCeilingFloor, &_|, ==_!=


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 32
# TOPICS: vectors, roundTruncCeilingFloor, &_|, ==_!=
#
# Write a command that shows the numbers in the vector nums that are both even
# and perfect squares. For example, if nums contains c(81, 6, 36, 10, 4, 25)
# then your answer should display 36  4
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x[sqrt(x)^2 == floor(x)  &  x%%2 == 0]    # ANSWER
```
:::

---

### QUESTION 33 TOPICS: vectors


```{r}
############################################################################################################.
# NOTE: The sort(x) function will return a vector with the values from the vector x in sorted order. 
#       The unique(x) function will return a copy of the vector x with all of the duplicate values removed. 
#       These functions will help you to answer the following questions
############################################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 33
# TOPICS: vectors
#
# Write a command that displays the 2nd lowest value in the vector x. If the lowest value appears
# more than once in the vector, then your code should display that value. For example, if x contains
# the numbers 7 3 3 7 3 5 20, the function should return the number 3. Even though 3 is the lowest
# value, it appears more than once, so in that sense, it is the 2nd lowest value.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(7,3,3,7,3,5,20)
sort(x)
sort(x)[2]            # ANSWER

# The following is just another example. It is NOT a new answer. 
x <- c(7,7,3,5,20) # ANOTHER EXAMPLE
sort(x)            # 
sort(x)[2]         # 
```
:::

---

### QUESTION 34 TOPICS: vectors


```{r}
############################################################################################################.
# NOTE: The sort(x) function will return a vector with the values from the vector x in sorted order. 
#       The unique(x) function will return a copy of the vector x with all of the duplicate values removed. 
#       These functions will help you to answer the following questions
############################################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 34
# TOPICS: vectors
#
# Write a command that displays the 2nd lowest value in the vector x. For this question, you
# should ignore duplicate values. For example, if x contains the numbers 7 3 3 7 3 5 20, the function
# should return the number 5. Even though 3 appears more than once, 5 is the 2nd lowest value.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# ONE ANSWER
x <- c(7,3,3,7,3,5,20)
unique(x)
sort(x)
sort(unique(x))
sort(unique(x))[2]            # ANSWER

# ANOTHER ANSWER
x <- c(7,3,3,7,3,5,20)
xWithoutLowestNumber = x[ x != min(x) ]  # ANSWER MULTIPLE LINES
min(xWithoutLowestNumber)                # ANSWER MULTIPLE LINES
```
:::

---

### QUESTION 35 TOPICS: vectors


```{r}
############################################################################################################.
# NOTE: The sort(x) function will return a vector with the values from the vector x in sorted order. 
#       The unique(x) function will return a copy of the vector x with all of the duplicate values removed. 
#       These functions will help you to answer the following questions
############################################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 35
# TOPICS: vectors
#
# Write a command that displays the n lowest values in the vector x where n is
# a variable that contains a positive integer. The values should be displayed
# in increasing order.
#
# For example if x contains c(67,2,40,5,2,99) and n contains the number 3 then
# the command should display the numbers 2 2 and 5
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
n <- 3
x <- c(67,2,40,5,2,99)
x
sort(x)
sort(x)[1:n]            # ANSWER
```
:::

---

### QUESTION 36 TOPICS: vectors


```{r}
############################################################################################################.
# NOTE: The sort(x) function will return a vector with the values from the vector x in sorted order. 
#       The unique(x) function will return a copy of the vector x with all of the duplicate values removed. 
#       These functions will help you to answer the following questions
############################################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 36
# TOPICS: vectors
#
# Same as previous question, but the numbers should be in decreasing order. 
# For example, given the values from the previous question, the answer should
# display 5 2 2. Hint, see the help for the sort command and focus on the
# "decreasing" argument to sort.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
n <- 3
x <- c(67,2,40,5,2,99)
x
sort(x)
sort(x)[1:n]
sort(sort(x)[1:n], decreasing=TRUE)   # ANSWER

# ALTERNATE ANSWER

n <- 3
x <- c(67,2,40,5,2,99)
x
x <- sort(x, decreasing=TRUE)  # ANOTHER ANSWER (MULTIPLE LINES)
x
(length(x)-n+1)
(length(x)-n+1):length(x)
x[(length(x)-n+1):length(x)]   # ANOTHER ANSWER (MULTIPLE LINES)
```
:::

---

### QUESTION 37 TOPICS: vectors


```{r}
############################################################################################################.
# NOTE: The sort(x) function will return a vector with the values from the vector x in sorted order. 
#       The unique(x) function will return a copy of the vector x with all of the duplicate values removed. 
#       These functions will help you to answer the following questions
############################################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 37
# TOPICS: vectors
#
# Same as previous question, but the numbers should be displayed in the same order
# as they appear in the original vector x. Given the same values from the previous
# question, the answer should display the numbers  2  5  2  in that order. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
n <- 3
x <- c(67,2,40,5,2,99)
sort(x)[n]
x[x <= sort(x)[n]]     # ANSWER
```
:::

---

### QUESTION 38 TOPICS: vectors


```{r}
############################################################################################.
# QUESTION_GROUP:
#
# (The next two questions are related to each other. I would have numbered them a,b,etc but 
#  I thought it would be better for each question to have its own number.)
##########################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 38
# TOPICS: vectors
#
# x and y each contain a single integer. nums is a vector that contains many numbers. 
# You are told that the value of x is guaranteed to be less than the value of y.
# Change the vector nums so that it only contains the values that are in the range
# between x and y (including x and y). For example, if x contains 3, y contains 7 and
# nums contains c(10,3,2,11,6,4,1) then your code should change the value of nums to a vector
# containing 3 6 and 4..
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# ONE POSSIBLE ANSWER
# The following answer is done in two different steps
x <- 3 # setup some example data
y <- 7 # setup some example data
nums <- c(10,3,2,11,6,4,1) # setup some example data

nums <- nums[nums>=x]      # ANSWER (MULTIPLE LINES) - nums will now contain all the values that
# are greater than or equal to x. Given the example data above, nums,
# will now contain c(10,3,11,6,4). The 2 and 1 were eliminated from nums.

nums                       # show the new values in nums
nums <- nums[nums<=y]      # ANSWER (MULTIPLE LINES) -  nums will now contain only those values 
# that are less than or equal to y. Given the example data above, nums
# will now contain c(3,6,4). The 10, and 11 were eliminated from nums.

nums                       # show the final value of nums


# ANOTHER POSSIBLE ANSWER  -  this answer is "shorter" and more "to the point". However, this answer
#                             requires the use of the & operator.
x <- 3   # setup some example data
y <- 7   # setup some example data
nums <- c(10,3,2,11,6,4,1)   # setup some example data

nums >= x      # given the example data above, this is a logical vector that contains TRUE for every value
# of nums that is >= x. Since nums contains c(10,3,2,11,6,4,1) and x contains the number 3,
# the logical vector will be c(TRUE,TRUE,FALSE,TRUE,TRUE,TRUE,FALSE)

nums <= y      # given the example data above, this is a logical vector that contains TRUE for every value
# of nums that is <= y. Since nums contains c(10,3,2,11,6,4,1) and y contains the number 7,
# the logical vector will be c(FALSE,TRUE,TRUE,FALSE,TRUE,TRUE,TRUE)
nums>=x & nums <= y   # This will combine the two logical vectors shown above using the & operator.
# That will result in a vector where there is a TRUE in the positions that had
# TRUE for both nums>=x and for num<=y (see above). In other words, the resulting
# logical vector will contain a TRUE in those positions that correspond to values
# in nums that are BOTH >=x AND <=y.
# 
#    as explained above, "nums >= x" is:    c(TRUE ,TRUE ,FALSE,TRUE ,TRUE ,TRUE ,FALSE)
#    as explained above, "nums <= y" is:    c(FALSE,TRUE ,TRUE ,FALSE,TRUE ,TRUE ,TRUE)
# 
# Therefore, "nums>=x & nums <= y" is:      c(FALSE,TRUE ,FALSE,FALSE,TRUE ,TRUE ,FALSE)

nums <- nums[nums>=x & nums <= y]  # ANSWER - "nums <-" simply means that we are going to replace 
#          the values in "nums" with the result of the code from the 
#          right hand side of the arrow.
#     
#          The code on the "Right Hand Side" (RHS) of the arrow is:
#          "nums[nums>=x & nums<=y]". The first "nums" on the RHS of 
#          the arrow indicates that we will be taking values from "nums".
#          the [brackets] that surround the rest of the code indicate 
#          that the code inside the [brackets] will be used to identify
#          which values from nums we will be "taking". As shown above
#          the code inside the brackets results in a logical vector
#          that contains a TRUE for each value that we want from nums. 
#          
#          Given the example data above, this code therefore becomes the 
#          equivalent of: 
#          nums <- nums[c(FALSE,TRUE ,FALSE,FALSE,TRUE ,TRUE ,FALSE)]
#          
#          This results in replacing nums with the 2nd 5th and 6th values
#          from nums (i.e. those positions that contains TRUE). Therefore
#          since "nums" originally contained c(10,3,2,11,6,4,1), the 
#          new value of "nums" will be c(3,6,4)
```
:::

---

### QUESTION 39 TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 39
# TOPICS: vectors
#
# Same as previous question, except for this question the value in x may be less than,
# greater than or equal to y. Hint: use the sort() function to isolate the smaller
# of x and y and the larger of x and y. Then proceed similarly to the way you answered
# the previous question. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# ONE ANSWER:
#
# To understand the answer to this question, you must first understand the answer to the 
# previous question. Please first review the answer to the previous question before continuing.
#
# The difference between this question and the previous question is that in the previous 
# question, you were told that the value of the variable "x" is a number that is LESS THAN
# the value in the variable "y". For example, in the previous question, it would be possible
# for x to be 3 and y to be 7 or x to b 2 and y to b 5, but it would NOT be possible for x
# to be 7 and y to be 3. HOWEVER, in this version of the question, x could be ANY number
# and y could be ANY number. 
#
# Therefore when you write the code to find the values in "nums" that are between x and y 
# you know that the numbers that you desire are all greater than x and less than y. Therefore
# in the previous question, you knew that nums>=x needed to be TRUE and nums<=y needed to be TRUE.
#
# However, in this question you DON'T know which number is less, x or y. Therefore
# if x is less than y then the code from the previous question would work. For example,
# if x is 3 and y is 7 then the expression   5>=3 & 5<=7   would result in a TRUE value.
#
# However, if y were less then x (e.g. x is 7 and y is 3) then "nums>=x & nums<=y" would
# result in a FALSE value for EVERY number. For example the number 5 is obviously between
# x=7 and y=3. However, the expression    5>=7 & 5<=3 would be FALSE.
#
# One approach to answer this question is given the values in x and y, to have your code
# "figure out" which value is higher and which is lower. You can store the lower value in a
# variable named "lower" and the higher value in a variable named "higher". Then you can simply use "lower"
# value as you used "x" in the previous question and use "higher" as you used "y" in the 
# previous question. This is exactly what is done by the code below 


x <- 7   # setup some example data
y <- 3   # setup some example data
nums <- c(10,3,2,11,6,4,1)   # setup some example data

sort(c(x,y))                # "c(x,y)" is a vector that contains the values of x and y. Given the
# example data above, since x is 7 and y is 3, c(x,y) is c(7,3). 
#
# The "sort" function takes a vector and sorts its contents. So
# sort(c(7,3)) results in the vector c(3,7)

lower <- sort(c(x,y))[1]  # As described above, given the example data above, the code:
# "sort(c(x,y))" will result in the vector c(3,7). The "[1]" that
# appears after "sort(c(x,y))" extracts the first value from the 
# sorted vector. Given the data above, the first value of c(3,7) is 3.
# This number, 3, is then assigned to the variable "lower".
lower
higher<- sort(c(x,y))[2]  # In a similar way, this code assigns the 2nd value from the sorted vector
# to the variable "higher". Specifically, given the example data, the value
# 7 will be assigned to the variable "higher".
# 
# Note that if the values in x and y were initially 3 and 7 instead of 7 and 3,
# we would still get the 3 being assigned to lower and 7 being assigned to higher.
# Now that we have variables "lower" and "higher", the answer is exactly
# the same as the answer to the previous question, except that instead of using
# x and y in the answer we will use "lower" and "higher".
higher
nums <- nums[nums>=lower]       # ANSWER (MULTIPLE LINES)
nums
nums <- nums[nums<=higher]      # ANSWER (MULTIPLE LINES)
nums

# ANOTHER POSSIBLE ANSWER  -  this answer is "shorter" and more "to the point". However, this answer
#                             requires the use of the & operator. See the explanation of the answer 
#                             to the previous question for more detail.
x <- 7
y <- 3
lower <- sort(c(x,y))[1]
higher<- sort(c(x,y))[2]
nums <- c(10,3,2,11,6,4,1)
nums <- nums[nums>=lower & nums <= higher]      # ANSWER
```
:::

---

### QUESTION 40 TOPICS: vectors


```{r}
############################################################################################.
# QUESTION_GROUP:
#
# The next few questions are related to each other.
# I would have numbered them a,b,etc but I thought it would be better
# for each question to have its own number.
############################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 40
# TOPICS: vectors
#
# nums is a vector that contains several numbers. n contains a single positive integer.
# Write code that "shifts" the values in nums over to the right "n" times. For example,
# if nums contains 10 20 30 40 50 60 70 80 90 100 and n is 2 then your command should
# reset nums to 90 100 10 20 30 40 50 60 70 80. If n were 3, your command should
# reset nums to 80 90 100 10 20 30 40 50 60 70
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# The following breaks up the answer into several lines so that it is easier to understand.
# The general idea of the solution below is to figure out which values from the original vector
# will be moved to the beginning of the vector. 

nums <- seq(10,100,by=10)   # setup some example data
nums
n <- 3                      # setup some example data
positionOfLastNumber <- length(nums)-n                     #ANSWER (MULTIPLE LINES) - This is the position
#   of the value from the original vector that 
#   will be in the last position of the answer.
#   For example if nums is:
#        c(10,20,30,40,50,60,70,80,90,100)
#   and n is 3 then the final answer will be
#        c(80,90,100, 10,20,30,40,50,60,70)
#   Therefore the 70 which is the last number 
#   in the answer, is in the 7th position in nums.

positionOfLastNumber
nums[positionOfLastNumber]
firstPiece <- nums[ (positionOfLastNumber+1):length(nums)] #ANSWER (MULTIPLE LINES) - this code captures
#   the first few numbers in the answer, which
#   are the last few numbers in nums. These numbers
#   start at the position after the position we 
#   calculated above and extend until the end of
#   the nums vector. 
# 
#   For example if nums is:
#        c(10,20,30,40,50,60,70,80,90,100)
#   and n is 3 then the final answer will be
#        c(80,90,100, 10,20,30,40,50,60,70)
#   Therefore the 70 which is the last number 
#   in the answer, is in the 7th position in nums.
#   However, the 80 (which is in the 7th+1 postion,
#   (i.e. the 8th position) in nums will be in the 
#   first position in the final answer. The rest
#   of the numbers in nums from the 80 through 100
#   will be in the first few positions of the answer.
# 
#   We use the variable "firstPiece" to capture
#   those numbers that will be in the first few 
#   positions in the answer.
firstPiece
lastPiece <- nums[1:(positionOfLastNumber)]                #ANSWER (MULTIPLE LINES) -  the last few numbers
#   in the answer will start from the first positon
#   in "nums" and extend until the postion that we
#   calculated above. We store those numbers
#   in the variable, "lastPiece".
lastPiece
nums <- c(firstPiece, lastPiece)                           #ANSWER (MULTIPLE LINES) - at this point we have
#   the first few numbers for the answer stored in 
#   the variable "firstPiece" and the last few
#   numbers from the answer stored in the variable
#   "lastPiece". All that remains to be done is 
#   to create a new vector with the numbers in that
#   order and assign the result back to the variable
#   "nums".
nums          


# Alternatively, you could write the answer in one line
nums <- seq(10,100,by=10)
nums
n <- 3
nums <- c(nums[ (length(nums)-n+1):length(nums)] , nums[1:(length(nums)-n)] )   # ALTERNATIVE ANSWER
nums          
```
:::

---

### QUESTION 41 TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 41
# TOPICS: vectors
#
# Same as previous question, but the numbers at the end of nums should be
# removed and the beginning of nums should be filled in with "n" zeros. For example,
# if nums contains 10 20 30 40 50 60 70 80 90 100 and n is 2 then your command should
# reset nums to 0 0 10 20 30 40 50 60 70 80. If n were 3, your command should
# reset nums to 0 0 0 10 20 30 40 50 60 70
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
nums <- seq(10,100,by=10)
nums
n <- 3
nums <- c(rep(0,n) , nums[1:(length(nums)-n)] )  # ANSWER
nums          
```
:::

---

### QUESTION 42 TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 42
# TOPICS: vectors
#
# Same as above, but replace the numbers at the beginning of nums with 1,2,3,etc.
# For example, if nums contains 10 20 30 40 50 60 70 80 90 100 and n is 2 then your
# command should reset nums to 1 2 10 20 30 40 50 60 70 80. If n were 3, your command
# should reset nums to 1 2 3 10 20 30 40 50 60 70
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
nums <- seq(10,100,by=10)
nums
n <- 3
nums <- c(1:n , nums[1:(length(nums)-n)] )   # ANSWER
nums          
```
:::

---

### QUESTION 43 TOPICS: vectors


```{r}
############################################################################################.
# QUESTION_GROUP: 
# (The next few questions are related to each other. I would have numbered them a,b,etc but 
#  I thought it would be better for each question to have its own number.)
############################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 43
# TOPICS: vectors
#
# The vector nums contains several numbers. Display the values of nums without the
# smallest numbers. For example, if nums contains c(30,5,20,5,30,15) then your command
# should display 30 20 30 15(none of the 5's are displayed since 5 is the smallest number).
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
nums <- c(30,5,20,5,30,15) 
min(nums)
nums != min(nums)
nums[nums != min(nums)]    # ANSWER
nums[nums > min(nums)]     # ALTERNATIVE ANSWER
```
:::

---

### QUESTION 44 TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 44
# TOPICS: vectors
#
# Same as the previous question, but do not display any of the smallest or the
# largest numbers. For example, if nums contains c(30,5,20,5,30,15) then your command
# should display 20 15 (none of the 5's are displayed since 5 is the smallest number and
# none of the 30's are displayed since 30 is the largest number).
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

##############################################################################################.
# END_OF_QUESTION_GROUP
##############################################################################################.
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# The best way to do this is with the "&" operator. If you haven't learned the "&" operator yet
# There are alternatives below
nums <- c(30,5,20,5,30,15) 
nums[nums != min(nums) & nums != max(nums)]   # ANSWER
nums[nums > min(nums) & nums < max(nums)]     # ALTERNATIVE ANSWER

# This version of the answer does NOT use the & operator, but requires two lines of code.
# These alternatives use the != (i.e. "not equal to") operator
nums <- c(30,5,20,5,30,15) 
temp <- nums[nums != min(nums)]  #ANSWER (MULTIPLE LINES)
temp[temp != max(nums)]          #ANSWER (MULTIPLE LINES)

# This version of the answer does NOT use the & operator, but requires two lines of code.
# These alternatives use the > and < operators instead of the != operator
nums <- c(30,5,20,5,30,15) 
temp <- nums[nums > min(nums)]  #ALTERNATIVE ANSWER (MULTIPLE LINES)
temp[temp < max(nums)]          #ALTERNATIVE ANSWER (MULTIPLE LINES)
```
:::

---

### QUESTION 45 TOPICS: vectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 45
# TOPICS: vectors
#
# x and y are two vectors whose lengths are the same. Merge the contents of
# vector y into vector x in the following way. If x1,x2,x3,x4,... represent the
# initial values in x and y1,y2,y3,y4,... represent the values in y, then after
# your commands run, the values in x should be x1,y1,x2,y2,x3,y3,x4,y4,...
#
# For example, if x contains the values 1 2 3 4 and y contains the values 11 22 33 44 
# then after your code runs the values in x should be 1 11 2 22 3 33 4 44. 
# (You may use more than one command to accomplish this result).
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(1,2,3,4)
x
y <- c(11,22,33,44)
y
seq(1,2*length(x),by=2)


x[seq(1,2*length(x),by=2)] <- x    # ANSWER (MULTIPLE LINES)

x
seq(2,2*length(y),by=2)


x[seq(2,2*length(y),by=2)] <- y    # ANSWER (MULTIPLE LINES)
x
```
:::

