---
format: html
---

## Topics: for loops

What is displayed by the following code?

```{r}

f = function(n){
  
  for (x in 1:n){
    cat(x * x, " ")
  }
  
  cat("\n")
  
  return(c(x, n))

}

# What does this display?
#result = f(4)

# What does this display?
#result
```

::: {.callout-note icon=false collapse="true"}
#### Click here for answer
```{r}
# What does this display?
result = f(4)

# What does this display?
result
```
:::

## Topics: while loops, cat vs return

What is displayed by the following code?

```{r}

f = function(n){
  
  x = 1
  while  (n >= 1){
    cat(x, " ")
    x = x * 2
    n = n - 1
  }
  
  cat("\n")
  
  return(c(x, n))

}

# What does this display?
#x = f(4)

# What does this display?
#x
```

::: {.callout-note icon=false collapse="true"}
#### Click here for answer
```{r}
# What does this display?
x = f(4)

# What does this display?
x
```
:::



## Topics: for loops

What is displayed by the following code?

```{r}
f = function(n){
  
  for (x in 1:n){
    cat("*** starting outer for loop again***\n")
    for (y in x:n){
      cat("x is ", x, "    ")
      cat("y is ", y, "\n")
    }
  }
  
  return( c(x,y) )
}

# What does this display?
#result = f(4)

# What does this display?
#result
```

::: {.callout-note icon=false collapse="true"}
#### Click here for answer
```{r}
# What does this display?
result = f(4)

# What does this display?
result
```
:::




## Topics: for loops, nested loops, if

What is displayed by the following code?

```{r}

drawSomething = function(size){
  
  if (!is.numeric(size) || trunc(size) != size || size < 0){
    stop("size must be a non-negative whole number")
  }
  
  for (row in 1:size){
    
    for (col in 1:size){
      
      if (col == row || col == size - row + 1){
        cat("@")
      } else {
        cat(".")
      }
        
    }
    
    cat("\n")
  }
  
  cat("\nHope you like the picture! :)\n")
    

}

# What do these display?
#drawSomething(3)
#drawSomething(5)
#drawSomething(1)
#drawSomething(0)  # be careful! This is tricky. Probably not what you'd expect.
```

::: {.callout-note icon=false collapse="true"}
#### Click here for answer
```{r}
# What do these display?
drawSomething(3)
drawSomething(5)
drawSomething(1)
drawSomething(0)  # be careful! This is tricky. Probably not what you'd expect.
```
:::




## Topics: for loops, nested loops, if

The only thing that changed are the lines that are highlighted 
in the code with "# THIS CHANGED".  
What will be displayed now?


```{r}
drawSomething = function(size){
  
  if (!is.numeric(size) || trunc(size) != size || size < 0){
    stop("size must be a non-negative whole number")
  }
  
  for (row in 1:size){
    
    for (col in 1:size){
      
      if (col == row || col == size - row + 1){  # THIS CHANGED
                                                 # THIS CHANGED
        if ( row %% 2 == 0)                      # THIS CHANGED
          cat("0")                               # THIS CHANGED
        else                                     # THIS CHANGED
          cat("X")                               # THIS CHANGED
                                                 # THIS CHANGED
      } else {
        cat(".")
        
      }
        
    }
    
    cat("\n")
  }
  
  cat("\nHope you like the picture! :)\n")
    

}

# What does this display?
#drawSomething(5)
```

::: {.callout-note icon=false collapse="true"}
#### Click here for answer
```{r}
# What does this display?
drawSomething(5)
```
:::




## Topics: for loops, nested loops, if

What is displayed by the following code?

```{r}
drawSomething = function(size){
  
  if (!is.numeric(size) || trunc(size) != size || size < 0){
    stop("size must be a non-negative whole number")
  }
  
  for(row in 1:size){

    if( row == 1 || row == size ){
      for (x in 1:size){
        cat("-")
      }
      cat("\n")
      
    } else {
      for (col in 1:size) {
        if( row == 1 || row == size ){
          for (x in 1:size){
            cat("-")
          }
          cat("\n")
          
        } else {
          if (col == 1){
            cat("|")
          } else if (col == size){
            cat("|\n")
          } else if (row == col) {
            cat("x")
          } else {
            cat(".")
          }
        }
      }
    }

  }
}

# What does this display?
#drawSomething(5)

# What does this display?
#drawSomething(10)
```


::: {.callout-note icon=false collapse="true"}
#### Click here for answer
```{r}
# What does this display?
drawSomething(5)

# What does this display?
drawSomething(10)
```
:::




