# Practice Questions

The following questions are related to the material for the midterm.
(I will try to keep adding questions to this list.)

1-6, 8-11, 14-23, 25-31, 33-42

::: {.content-hidden}
# NOTE:
#
# [The following notes are for Prof. Rosenthal (and anyone else who will be
# editing this file) to aid him in modifying this file.]
#
# The original versions of this file were created in MS Word. Word files are
# able to be printed much more concisely than this format of the file (in R) since
# an entire question is in one paragraph that automatically wraps to the correct
# width. By contrast, in this R code version of the file I have to break up a 
# single question into multiple lines so that the #commentCharacter comes out at 
# the beginning of the line. This can make the file much longer and more difficult
# to print. Also in Word, it's very easy to use a tiny font to print. In an
# R file - it seems to be less easy to do so. 
#
# Nevertheless, I converted the MSWord files to .R files. This makes it much easier
# to maintain this file (with solutions) and convert it to an equivalent file
# with just the questions by simply taking out the answers/solutions.
#
# The following information can be used to automatically generate the 
# question file (that does not contain the answers.)
#    - All questions start with a number and a period. 
#    - The entire question in is a single comment
#    - The answers start below that and continue to the next question.
:::

::: {.content-hidden}
# TOPICS

==_!=
%/%_%%
&_|
&&_||
vectors
recycling 
vectorArithmetic
vectorAssignment
indexingVectors
indexingWithLogicalValues
indexingWithNumbers
indexingWithLogicalValues 
c
sum_mean
min_max
round_trunc_ceiling_floor
rep_seq_colonOperator
sqrt
length
runif
namedVectors
orderOfOperations
parallelVectors
userDefinedFunctions

dataframes
loops
lists
sciNotation
Inf


## indexing with positive numbers
## indexing with negative numbers
## indexing with logical values
## length
## round
## R help pages
## seq
## rep
## :
## c
## named vectors
## vector assignment
:::


## About these questions/answers

```{r}
###################################################################################.
###################################################################################.
###################################################################################.
#####.
#####                             IMPORTANT:
#####.
#####                 READ THE INFORMATION BELOW
#####    BEFORE READING THE REST OF THIS DOCUMENT!
#####.
###################################################################################.
###################################################################################.
###################################################################################.

###########################################################################################################.
#
# R QUESTIONS AND ANSWERS
#
# This file contains R questions and answers. This file is actually a runnable R script.
# The questions are in comments. The answers are in runnable code. To better understand the answers,
# I highly recommend that you copy the questions and answers into an R script and run the commands 
# one at a time and view the results.
#
# The actual answer is on the line containing the comment #ANSWER. All other lines are there to
# demonstrate the answer working on actual data or to demonstrate portions of the whole answer so
# that the whole answer is more easily understood. If the actual required code for an answer takes
# up several lines then the comment will say #ANSWER (MULTIPLE LINES). 
#
# If different possible answers are shown, then the different answers will be displayed with the comment
# #ALTERNATE ANSWER or #ANOTHER ANSWER.  Note that it is very possible that there are valid answers
# other than the ones shown below.
#
###########################################################################################################.
```

## Vectors

### QUESTION 1 *(vectors, rep/seq/:/etc)*

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION:
#
# Write an R command that displays all the even numbers between 100 and 1000 (including 100 and 1000) 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
seq(100,1000,by=2)  #<<ANSWER>>
```
:::

---

### QUESTION 2 *(vectors , rep/seq/:/etc)*

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION:
#
# Write an R command that displays the values:   10, 13, 16, 19, 22, etc. There
# should be 500 numbers in all.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
seq(10,by=3,length.out=500)  #ANSWER
```
:::

---

### QUESTION 3. *(vectors , rep/seq/:/etc)*

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION
#
# Write an R command that displays alternating values of positive and
# negatives of 1,2,3 ie.:  1,2,3,-1,-2,-3,1,2,3,-1,-2,-3,1,2,3 ... etc.
# There should be a total of 60 numbers in all.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
rep(c(1:3,-1:-3),length.out=60)   # ANSWER

rep(c(1,2,3,-1,-2,-3),length.out=60)   # ANOTHER ANSWER
```
:::

---

### QUESTION 4. *(vectors , recycling)*

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 4    TOPICS: vectors , recycling  ####
# 
# Write an R command that displays all the powers of 2 from 
# 2 to the power of minus 20 (i.e. -20) until
# 2 to the power of positive 20 (i.e. +20)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
-20:20      # this gives the numbers -20 -19 -18 ... -2 -1 0 1 2 ... 18 19 20

2^(-20:20)  # ANSWER

2^-20:20    # NOTE: This is WRONG because order of operations makes this work like (2^-20):20
```
:::

---

### QUESTION 5 *(vectors , indexingWithLogicalValues)*

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 
#
# Write an R command that displays every entry in the vector x whose value is 10 or more.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(3,10,11,2,-5,100,21)  # setup some example data
x
x >= 10       # This gives TRUE/FALSE values - there will be one 
              # TRUE or FALSE for each value in x. Given the data above,
              # this will be: c(FALSE,TRUE,TRUE,FALSE,FALSE,TRUE,TRUE)

x[x>=10]      # ANSWER - The first "x" outside the [brackets] is the vector
              #          whose values are going to be displayed. The 
              #          "x>=10" inside the [brackets] is used to generate
              #          a logical vector (i.e. TRUE/FALSE values) to indicate
              #          which values will be displayed. Only those values
              #          from x whose positions correspond to a TRUE value
              #          in the logical vector, will be displayed. Given the
              #          sample data above, the 2nd, 3rd, 5th and 7th values
              #          will be displayed, i.e. 10,11,100,21
```
:::

---

### QUESTION 6    TOPICS: vectors , indexingWithLogicalValues ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 6    TOPICS: vectors , indexingWithLogicalValues ####
#
# x and y are two numeric vectors whose lengths are equal.
# Show every value from the vector, x, that is less than the corresponding value in the vector y
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(3,10,11,2,-5,100,21)  # setup some example data 
y <- c(10,1,11,15,5,9,999)   # setup some example data
length(x)
length(y)
x < y         # This gives TRUE/FALSE values - there will be one 
              # TRUE or FALSE for each value in x. Given the data above,
              # this will be: c(FALSE,TRUE,TRUE,FALSE,FALSE,TRUE,TRUE)

x[x<y]     # ANSWER
y[x<y]     # FYI - this wasn't asked but these are the corresponding
           # values from y that are greater than those in x
```
:::

---

### QUESTION 7    TOPICS: vectors , round , runif , RHelpPages  ####


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 7    TOPICS: vectors , round , runif , RHelpPages  ####
#
# SOME BACKGROUND - runif function
#
#     The runif function returns a numeric vector that contains "random numbers
#     with a uniform distribution". The R help page for runif shows that the
#     arguments and default values for those arguments are:
#         runif(n, min=0, max=1)     
#     where
#         n is the number of numbers to generate
#         min is the smallest value that a number could have
#         max is the largest value that a number could have. 
#
#     EXAMPLES: 
#  
#     > # return a vector of 3 random numbers whose values fall between 0 and 10
#     > runif(3,0,10)   
#     [1] 3.785233 9.672244 2.857430
# 
#     > # return a vector of 10 random numbers whose values fall between -5 and +5
#     > runif(10,-5,5)
#     [1] -1.84032071  4.28685602  2.59656498 -1.09880205 -3.90202237 -1.89618831  3.37193245
#     [8]  0.08757248 -1.10848900 -1.90661883
# 
#     > # return a vector of 5 random numbers. The min and max are not specified so 
#     > # use the default values of min (ie. 0) and max (i.e. 1)
#     > runif(5)
#     [1] 0.4098594 0.8771700 0.5610475 0.8928674 0.4863841 
#
#   WHAT YOU HAVE TO DO:
#
#     Generate 10 random numbers that are between 50 and 51. The numbers should have
#     2 values after the decimal point. (HINT: use the   runif   and the   round
#     functions). Note that every time you run this command the actual numbers will
#     be different.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
runif(10,50,51)                  # generate 10 random numbers between 50 and 51

round(runif(10,50,51),digits=2)  # ANSWER 
```
:::

---

### QUESTION 8    TOPICS: vectors , c ####


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 8    TOPICS: vectors , c ####
#
# Insert the number 5 in the beginning of the vector named x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(7,8,2,1)  # start with some data for x (this could be ANY data)
x <- c(5,x)      # ANSWER -  x is now c(5,7,8,2,1)
x             

x[1] <- 5    # WRONG: This will change the first position to 5
# given the example data above the 7 will be overwritten with the 5
# and x will be c(5,8,2,1)
```
:::

---

### QUESTION 9   TOPICS: vectors , c , length  ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 9   TOPICS: vectors , c , length  ####
#
# Insert the number 6 at the end of the vector named x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# ONE WAY
x <- c(7,8,2,1)    # setup some example data
x <- c(x,6)    # ANSWER
x


# ANOTHER WAY
x <- c(7,8,2,1)    # setup some example data
x[length(x)+1] = 6   # ANOTHER ANSWER
x
```
:::

---

### QUESTION 10   TOPICS: vectors , c , indexingWithNumbers

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 10   TOPICS: vectors , c , indexingWithNumbers
#
# x is a vector that contains 11 or more values.
# Insert the number 7 between the 10th and 11th values in the vector named x.
#
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- seq(10,100,by=10)    # setup some example data - x is now:
x <- seq(10,150,by=10)    # setup some example data - x is now:
#   c(10,20,30,40,50,60,70,80,90,100,110,120,130,140,150)
x
x <- c(x[1:10],7,x[11:length(x)])  # ANSWER 
x
#   x is now :
#   c(10,20,30,40,50,60,70,80,90,100,7,110,120,130,140,150)
#
#   The "length(x)" is necessary so that this answer
#   will work no matter how many values the vector, x, contains.
#   the data above was just and example, but the command should 
#   now matter what values x contains. 

x
```
:::

---

### QUESTION 11   TOPICS: vectors , c , indexingWithNumbers ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 11   TOPICS: vectors , c , indexingWithNumbers ####
#
# The vector x contains 10 values. Swap the first 5 and 
#     the last 5 values in x. For example, if x contains the 
#     values 1,2,3,4,5,6,7,8,9,10  then after your command runs, 
#     x should contain the values 6,7,8,9,10,1,2,3,4,5
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- seq(10,100,by=10) # setup example data x contains c(10,20,30,40,50,60,70,80,90,100)

x
x <- c(x[6:10],x[1:5]) # ANSWER -  x now contains c(60,70,80,90,100,10,20,30,40,50)
x
```
:::

---

### QUESTION 12   TOPICS: namedVectors ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 12   TOPICS: namedVectors ####
#
# Grades is a named vector that contains test scores.
#     The names of the entries are the names of the students.
#     Exchange the value of Tom's grade with Joe's grade where
#     "tom" and "joe" are the names of the values that contain
#     the grades for Tom and Joe.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
grades <- c(90,80,70,100,85)                        # setup some example data: set the values in the vector
names(grades) <- c("sue","tom","bill","ann","joe")  # setup some example data: set the names of the values
names(grades)                                       # this displays just the names: "sue","tom",etc.
grades                                              # this displays the values:  90,80,70,100,85
grades[c("tom","joe")]                              # this displays tom's grade and joe's grade: 80  85
grades[c("joe","tom")]                              # this displays joe's grade and tom's grade: 85  80

grades[c("tom","joe")] <- grades[c("joe","tom")]  # ANSWER: this puts the value of 
# Joe's grade into the position for Tom's grade and
# Tom's grade into the position for Joe's grade.
grades                                              # show the new vector
```
:::

---

### QUESTION 13   TOPICS: namedVectors , indexingWithLogicalValues ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 13   TOPICS: namedVectors , indexingWithLogicalValues ####
#
# Grades is a named vector that contains test scores. The names 
# of the entries are the names of the students. Change susan's name to sue
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
grades <- c(90,80,70,100,85)
names(grades) <- c("susan","tom","bill","ann","joe")
grades
names(grades)=="susan"   # Shows TRUE/FALSE values
names(grades) [ names(grades)=="susan" ]  # identify item in the names vector that contains susan

# The following line is the ONLY line that is required for the answer.
# The code puts the value "sue" into the position of the vector, names(grades)
# that currently contains the value "susan"

names(grades) [ names(grades)=="susan" ] <- "sue"       # ANSWER 

grades
```
:::

---

### QUESTION 14   TOPICS: vectors , indexingWithNumbers , vectorAssignment ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 14   TOPICS: vectors , indexingWithNumbers , vectorAssignment ####
#
# Exchange the 4th and 7th value in the vector nums.
#
#      a.  Do this using 3 different commands. Hint, you may create new variables. 
#          (Hint, you may create new variables. For example, I'll start you off:
#              fourth <- x[4]
#              seventh <- x[7]
#              # fill in the rest of the R code to answer the question.
#
#      b. Do this using only a single R command
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
#14a
nums <- seq(10,100,by=10)
x <- nums[7]        # ANSWER (MULTIPLE LINES)
nums[7] <- nums[4]  # ANSWER (MULTIPLE LINES)
nums[4] <- x        # ANSWER (MULTIPLE LINES)
nums

#14b
nums <- seq(10,100,by=10)
nums[c(4,7)] <- nums[c(7,4)]   # ANSWER
nums
```
:::

---

### QUESTION 15   TOPICS: vectors , indexingWithNumbers , length

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 15   TOPICS: vectors , indexingWithNumbers , length
#
# Display the last value from the vector x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(10,20,30,40,50)
x[length(x)]    # ANSWER
```
:::

---

### QUESTION 16   TOPICS: vectors , indexingWithNumbers , length

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 16   TOPICS: vectors , indexingWithNumbers , length
#
# Display the 2nd to last value in x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(10,20,30,40,50)
x[length(x)-1]          # ANSWER
```
:::

---

### QUESTION 17   TOPICS: vectors , indexingWithNumbers , mean , orderOfOperations ####

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 17   TOPICS: vectors , indexingWithNumbers , mean , orderOfOperations ####
#
# Display the average of the last 5 values in x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(1,5,2,3,1,20,4,10,6,50)
mean(x[(length(x)-4):length(x)])   # ANSWER

# Note, in the answer above, notice the parentheses around (length(x)-4).
# These parentheses are REQUIRED due to order of operations. Without the parentheses
# the command, length(x)-4:length(x) would actually be interpreted by R as 
# length(x)-(4:length(x)) which would give the WRONG result. For more information, 
# see the "recipe" about "order of operations in R" in the "R Cookbook" published by O'Reilly
```
:::

---

### QUESTION 18a   TOPICS: vectors , vectorArithmetic 

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 18a
# TOPICS: vectors, vectorArithmetic, seqColon, indexingWithLogicalValues, length, sciNotation, Inf
#
# Inf is used for "Infinity"
# 
# R can only work with numbers up to a certain size. Any value that is more
# than the maximum is returned as Inf (which stands for Infinity).
#
# Example:
#
# SCIENTIFIC NOTATION
#
# By default R displays very large and very small numbers using a mathematical
# notation called "scientific notation". "Scientific notation" is just a
# fancy way of writing very large and very small numbers using very little
# typing.
#
# To understand "scientific notation", let's start with a simple example.
# 1e+3 is the scientific notation way to write the number 1000.
# 1e+3 can be understood as 1*10^3 (as explained in the next paragraph)
#
# People who are not familiar with scientific notation might read
# "1e+3" as "one e plus 3". However, that is NOT the right way to read it.
# Instead, whenever you see "e" in scientific notation, read it as though
# instead of "e" it said "... times 10 to the power of ...", so "1e+3" should
# be read as "one times 10 to the power of positive 3"
# i.e., "1e+3" is the same as 1*10^3. 
# ( By the way, the letter "e" is used because, "e" is short for "exponent". )
#
# In a similar way, 2.3e+12 is the scientific notation way of writing the
# number 2300000000000 which is the same as "2.3 times 10 to the power of 12"
# i.e. 2.3*10^12
#
# To see that R "understands" scientific notation, just type 1e+3 into R.
# Because 1e+3 is the same as 1000, R just displays 1000. For example:
#
#   > 1e+3      # this is scientific notation for 1000.
#   [1] 1000
#
# By default if a number is small enough then R displays it in it's "regular"
# form (i.e. NOT in scientific notation). However, if a number is very large or
# extremely small, then R displays the number in scientific notation:
#
#   > # "regular numbers" get displayed normally
#   > 1000      
#   [1] 1000
#
#   > # very large numbers get displayed in scientific notation (with a positive exponent)
#   > 1000000000000  
#   [1] 1e+12
#
#   > # extremely small numbers get displayed in scientific notation (with a negative exponent)
#   > 0.0000000000001   
#   [1] 1e-13
#
# -----------------------------------------------------------------------------
# Write code that answers the following question:
#
#    What is the highest power of 4 that is less than one million?
#
# Your code should produce the answer 9. 
#
# The answer is 9 since 4^9=262,144 which is less than 1 million
# but 4^10=1,048,576 is more than one million.
# 
# Write R code to calculate this answer (i.e. 9) automatically.
# Your code should work for any base that is between 2 and 1000 (because 1000 is the
# square root of one million), not only for a base of 4. For example, if instead of
# the number 4, your code used the variable "n", your code should print the
# highest power of "n" that is less than one million. You may write the answer
# using several lines of R code. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
n <- 4               # ANSWER (MULTIPLE LINES)
n2 <- n^(1:1000)     # ANSWER (MULTIPLE LINES)
n2
n2 < 1000000
n2 <- n2[n2<1000000] # ANSWER (MULTIPLE LINES)
n2
length(n2)           # ANSWER (MULTIPLE LINES) (this will display the number 9)
```
:::

---

### QUESTION 18b    TOPICS: vectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 18b
# TOPICS: vectors, vectorArithmetic, seqColon, indexingWithLogicalValues, length
#
# Same as previous question, but this time n contains a "base" as before and
# x contains the maximum value (In the previous question this was one million.
# In this question it can be any number that is contained in x.) The question
# now is what is the maximum whole-number power that n can be raised to whose
# result is less than x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- 200
n <- 3               # ANSWER (MULTIPLE LINES)
n2 <- n^(1:sqrt(x))  # ANSWER (MULTIPLE LINES)
n2 <- n2[n2<x]       # ANSWER (MULTIPLE LINES)
length(n2)           # ANSWER (MULTIPLE LINES)
```
:::

---

### QUESTION 19    TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 19
# TOPICS: vectors, seqColon, indexingWithNumbers, vectorArithmetic
#
# nums is a vector that contains 100 numbers. 
#
# Use nums to create a vector that contains 50 numbers, as described below. 
# The first value should be the sum of the first two numbers from nums. 
# The 2nd value should be the sum of the 3rd and 4th values from nums. 
# For example, if nums contains 1,2,3,4,5,6 ... 99,100, then your answer 
# should contain 3,7,11 ... 199. 
#
# For example, if nums contains
#     3,6,2,8,1,-2, ... ninety-one-other-numbers ... 8,5,-5,2
#
# then your answer should contain
#     9,10,-1, ... forty-five-other-numbers ... 13,-3. 
#
# You may write the answer using several lines of R code.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
nums <- 1:100
firstThirdFifthEtc <- nums[seq(1,100,by=2)]   # ANSWER (MULTIPLE LINES)
secondFourthSixthEtc <- nums[seq(2,100,by=2)]     # ANSWER (MULTIPLE LINES)
answer <- firstThirdFifthEtc + secondFourthSixthEtc  # ANSWER (MULTIPLE LINES)
length(answer)
firstThirdFifthEtc
secondFourthSixthEtc
answer
```
:::

---

### QUESTION 20    TOPICS: vectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 20
# TOPICS: vectors, ==and!=, indexingWithLogicalValues, seqColon
#
# x and y are vectors that each contain 5 numbers. 
#
# PART A:   Show the values from x whose corresponding values in y are the same.
# PART B:   Show the position numbers that contain the same values in x and in y.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
#20a
x <- c(30,50,-20,10,40) # setup some example data
y <- c(99,50,-20,88,40) # setup some example data
x == y                  # This generates a vector that contains five TRUE/FALSE values

x[x==y]                 # ANSWER - "x==y" generates a logical vector (i.e. TRUE/FALSE values).
#          There will be one TRUE or FALSE value for each pair of values
#          in x and y. The "x" outside of the [brackets] indicates the 
#          vector whose values will be displayed. The values from x
#          that occupy the same positions as the TRUE values from "x==y"
#          will be displayed. The values from x that occupy the same positions
#          as the FALSE values from "x==y" will NOT be displayed.


#20b
x <- c(30,50,-20,10,40) # setup some example data
y <- c(99,50,-20,88,40) # setup some example data

(1:length(x))[x==y]     # ANSWER - NOTE: "x==y" generates a logical vector (i.e. TRUE/FALSE values).

#          There will be one TRUE or FALSE value for each pair of values
#          in x and y.
# 
#          The code: "(1:length(x))" that appears outside the [brackets]
#          generates a vector whose values are:
# 
#                   1  2  3  ... etc (up to the length of x).                        # 
#          Given the data above, x contains 5 values. Therefore 
#          the value of (1:length(x))will be:    1  2  3  4  5
# 
#          The code "[x==y]" is used to specify which of the values from the vector
#          should be displayed. The values from "(1:length(x))"
#          that occupy the same positions as the TRUE values from "x==y"
#          will be displayed. The values from x that occupy the same positions
#          as the FALSE values from "x==y" will NOT be displayed.
```
:::
  
---

### QUESTION 21 TOPICS: vectors, sumMean, indexingWithLogicalValues

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 21
# TOPICS: vectors, sumMean, indexingWithLogicalValues
#
# Show all values from the vector x that are greater than the average of the values in x. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(30,50,-20,10,40)
mean(x)
x[x > mean(x)]    # ANSWER
```
:::

---

### QUESTION 22   TOPICS: vectors, sumMean, roundTruncCeilingFloor


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 22
# TOPICS: vectors, sumMean, roundTruncCeilingFloor
#
# The sd() function returns the standard deviation of a vector.
#    Show all of the numbers in the vector x that are at least 1.5 standard deviations
#    away from the mean of x. You can show the numbers in any order you like. They do not
#    have to be in the same order that they appear in the vector x.
#
# HINT:  Standard Deviation is a concept that is covered in statistics. If you didn't take
# stats yet, I can understand why you're confused.  Bottom line, is that sd(c(3,10,76,32,23,31))
# will return a SINGLE number that is called the "standard deviation" of the values
# 3,10,76,32,23,31. sd(c(3,10,76,32,23,31) is 25.67. The mean of those same values
# is 29.16.  Therefore, numbers that are at least 1.5 standard deviations away from the mean are 
#    greater than 29.16 + 1.5 * 25.67    = 67.67
#    or less than 29.16 - 1.5 * 25.67    =  -9.34
# Therefore, if x contained the numbers shown, then you should only display the number 76 as
# that is the only number that is outside of the specified range.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- floor(runif(50,1,100))
x
sd(x)
mean(x)
mean(x) - 1.5 * sd(x)
mean(x) + 1.5 * sd(x)
x[x <= mean(x) - 1.5 * sd(x)]
x[x >= mean(x) + 1.5 * sd(x)]
c ( x[x <= mean(x) - 1.5 * sd(x)]  ,  x[x >= mean(x) + 1.5 * sd(x)]  )    # ANSWER
```
:::

---

### QUESTION 23   TOPICS: vectors, parallelVectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 23
# TOPICS: vectors, parallelVectors
#
# The following variables are parallel vectors that contain info
# about 4 different rectangles.
#
#   colors  <- c("red","blue","green","yellow)
#   lengths <- c(4,5,6,7)
#   widths  <- c(5,6,2,4)
#            
# In other words:
#   the 1st rectangle is "red",    it's length is 4 and it's width is 5
#
#   the 2nd rectangle is "blue",   it's length is 5 and it's width is 6
#
#   the 3rd rectangle is "green",  it's length is 6 and it's width is 2
#
#   the 4th rectangle is "yellow", it's length is 7 and it's width is 4
#
# QUESTIONS:
#
#    The actual numbers and colors shown above are just an example.
#    Your answers must work even if the vectors contain different 
#    numbers and colors and even if the vectors contains fewer or more values.
#
#    a. Show the AREAS of all the rectangles.
#
#            Given the example data above, 
#            your code should evaluate to: 20  30  12 28
#
#    b. Show the COLORS of the rectangles whose areas are at least 25.
# 
#            Given the example data above, 
#            your code should evaluate to: "blue"  "yellow"
#
#    c. Show the COLORS of the rectangles 
#       whose length is at least 5
#       or whose width is at least 5.
#
#           Given the example data above,
#           your code should evaluate to: "red"  blue"  "green" "yellow"
#
#    d. Show the COLORS of the rectangles 
#       whose length and width are both at least 5.
#
#           Given the example data above,
#           your code should evaluate to: "blue"
# 
#    e. Show the COLORS of the rectangles whose area is at least 25.
#
#           Given the example data above,
#           your code should evaluate to: "blue"  "yellow"
# 
#    f. Show the LENGTHS of the blue rectangles whose area is at least 25
#
#           Given the example data above,
#           your code should evaluate to: 5
# 
#    g. Show the LENGTHS of the rectangles whose area is at least 25 
#       and whose color is any one of blue, green or yellow.
#
#           Given the example data above,
#           your code should evaluate to: 5 7
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# 23a.

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

lengths * widths                          # ANSWER



# 23b. 

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

colors[(lengths * widths) >= 25]           # ANSWER




# 23c.

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

colors[lengths>=5 | widths>=5]           # ANSWER


# 23d.

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

colors[lengths>=5 & widths>=5]           # ANSWER


#23e.
# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

colors[lengths*widths>=25]           # ANSWER



# 23f.

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

lengths[colors=="blue" & lengths*widths>=25]           # ANSWER



# 23g. Show the lengths of the rectangles whose area is at least 25 and whose color
#       is any one of blue, green or yellow.

# setup some sample data as described in the question
lengths <- c(4,5,6,7)
widths <- c(5,6,2,4)
colors <- c("red","blue","green","yellow")

# ONE ANSWER: the parentheses are necessary due to the order of operations of & before |
lengths[lengths*widths>=25 & (colors=="blue" | colors=="green" | colors=="yellow") ]    # ANSWER

# ANOTHER ANSWER: this uses the %in% operator
lengths[lengths*widths>=25 & colors %in% c("blue","green","yellow") ]    # ANOTHER ANSWER
```
:::

---

### QUESTION 24   TOPICS: vectors, parallelVectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24
# TOPICS: vectors, parallelVectors
#
# The vector "roomDimensions" is used to store the lengths, widths and
# heights of several rooms in the following format:
# lengthRoom1,widthRoom1,heightRoom1,lengthRoom2,widthRoom2,heightRoom2, etc. 
#
# FOR EXAMPLE, the following code :
#
#         > roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)
#
# represents information about four different rooms with the following dimensions. 
#
#          first room:  length is 10, width is 4, height is 10
#          second room: length is 11, width is 5  height is 9
#          third room: length is 12, width is 5, height is 10.5
#          fourth room: length is 8,  width is 5, height is 9.5
#
# Different numbers may be assigned to roomDimensions, but the structure is as 
# described above. 
#
#    ANSWER PARTS a,b,c,d,e SHOWN BELOW
#      The answer to (a) should display:   4
#      The answer to (b) should display:   11  5  9
#      The answer to (c) should display:   40  55  60  40
#      The answer to (d) should display:   2  3
#      The answer to (e) should display:   9  10
# 
#      NOTE that your code for (a),(b),(c),(d) and (e)  should work no
#      matter what values are placed in the variable, roomDimensions
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

---

#### QUESTION 24a (see the intro to question 24 above)   # TOPICS: vectors, length


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24a (see the intro to question 24 above)
# TOPICS: vectors, length
#
# Write R code to display how many "rooms" are represented in the vector,
# i.e. 1/3 of the number of values in the vector.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# set the roomDimensions vector to some sample data as described in the question
#
#   The first  room is 10ft long, 4ft wide, 10ft tall (ie. the first three numbers in the vector)
#   The second room is 11ft long, 5ft wide, 12ft tall (ie. the next three numbers in the vector)
#   The third  room is 10ft long, 4ft wide, 10ft tall (ie. the next three numbers in the vector)
#   The fourth room is 11ft long, 5ft wide, 12ft tall (ie. the next three numbers in the vector)
roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)

# The number of rooms is 1/3 the number of numbers (since each room is described with 3  numbers)
length(roomDimensions) / 3   # ANSWER
```
:::

---

#### QUESTION 24b (see the intro to question 24 above)   TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24b (see the intro to question 24 above)
# TOPICS: vectors
#
#    Write R code to display the length, width and height of the nth room 
#    where n is a variable that contains a number. For example, if "n" contains 
#    the number 4 then your code should display the 3 numbers that represent the 
#    length, width and height of the 4th room.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# setup some sample data (see the question above)
roomDimensions <- c(10,  4,  10,  11,  5,  9,  12,  5,  10.5,  8,  5,  9.5)

n <- 4    # This is the number of the room we want (i.e. 1 for 1st room, 2 for 2nd room, etc)

# Display the dimensions for just the nth room
#
# To explain the following answer, it helps to think of a particular example.
# Suppose we were trying to get the dimensions for the 4th room ...
# ... keep in mind, that the first number for the 4th room appears right after the last number for the 3rd room.
# ... since there are 3 numbers for each room there are 9 numbers before the first number for the 4th room.
# ... i.e. there are 3*3 = 9 numbers for the first 3 rooms so the 4th room starts at position 10.
# ... to generalize, the first number for the nth room starts at (n-1)*3+1. In the example I just mentioned
# ... n is 4, so (n-1) is 3, so (n-1)*3 is 9 so the first number for the 4th room starts at position
# ... (n-1)*3+1 which is position 10.


roomDimensions[c(10,11,12)]

n=4
roomDimensions[c(10,11,n*3)]


roomDimensions[ c( (n-1)*3+1 , (n-1)*3+2 , (n-1)*3+3 ) ] # ANSWER


# Another way to think of this problem is that the last number for the 4th room is
# at position 4*3 i.e. the 12th number. Since each room as 3 numbers, the first number
# for the 4th room is at position 4*3-1 and the second number for the 4th room is at position 4*3-2.
# Therefore the 3 numbers for the 4th room are at positions 4*3-2 (i.e. 10) , 4*3-1 (i.e. 11) and 4*3 (i.e. 12)
# To generalize, the 3 numbers for the nth room are at positions n*3-2 , n*3-1 and n*3 

roomDimensions[ c(n*3-2, n*3-1, n*3) ]                   # ANOTHER ANSWER
```
:::

---

#### QUESTION 24c (see the intro to question 24 above)  # TOPICS: vectors, parallelVectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24c (see the intro to question 24 above)
# TOPICS: vectors, parallelVectors
#
# Write R code to display the number of square feet of floor space in each room.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# setup some sample data (see the question above)
roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)        

# Get a vector with just the lengths of each room (i.e. 10,11,12,8)
lengths <- roomDimensions[ seq(1,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES)   10 11 12 8
lengths

# Get a vector with just the widths of each room (i.e. 4,5,5,5)
widths <- roomDimensions[ seq(2,length(roomDimensions),by=3)  ] # ANSWER (MULTIPLE LINES) 4 5 5 5
widths

# multiply the lengths and widths to get the square feet of floor space
lengths * widths                                                # ANSWER (MULTIPLE LINES) 40 55 60 40
```
:::

---

#### QUESTION 24d (see the intro to question 24 above) TOPICS: vectors, parallelVectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24d (see the intro to question 24 above)
# TOPICS: vectors, parallelVectors
#
# Write R code to display the room numbers of those rooms that have 
# at least 50 square feet of floor space.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# setup some sample data (see the question)
roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)        

# extract just the lengths into a new vector (10 11 9 10.5)
lengths <- roomDimensions[ seq(1,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES) 10 11 9 10.5
lengths

# extract just the widths into a new vector (4 5 5 5)
widths <- roomDimensions[ seq(2,length(roomDimensions),by=3) ]  # ANSWER (MULTIPLE LINES) 4 5 5 5
widths

# create a vector with the room numbers, i.e. 1 2 3 4
roomNumbers <- 1:(length(roomDimensions)/3)                     # ANSWER (MULTIPLE LINES) 1 2 3 4
roomNumbers

# display the room numbers for those rooms whose length*width is at least 50
roomNumbers[lengths * widths >= 50]                             # ANSWER (MULTIPLE LINES)  2 3
```
:::

---

#### QUESTION 24e (see the intro to question 24 above)  TOPICS: vectors, parallelVectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 24e (see the intro to question 24 above)
# TOPICS: vectors, parallelVectors
#
# Write R code to display the heights of those rooms that have 
# at least 50 square feet of floor space.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
######################.
# ONE POSSIBLE ANSWER
######################.
# setup some sample data (see the question)
roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)        

# extract just the lengths into a new vector (10 11 9 10.5)
lengths <- roomDimensions[ seq(1,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES) 10 11 9 10.5
lengths

# extract just the widths into a new vector (4 5 5 5)
widths <- roomDimensions[ seq(2,length(roomDimensions),by=3) ]  # ANSWER (MULTIPLE LINES) 4 5 5 5
widths

# extract just the heights into a new vector (10 9 10.5 9.5)
heights <- roomDimensions[ seq(3,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES)
heights

# get the heights for just the rooms whose area is at least 50 (final answer)
heights[lengths * widths >= 50]                                # FINAL ANSWER (MULTIPLE LINES)  9 10



###########################.
# ANOTHER POSSIBLE ANSWER
###########################.

# setup some sample data (see the question)
roomDimensions <- c(10,4,10,11,5,9,12,5,10.5,8,5,9.5)        

# extract just the lengths into a new vector (10 11 9 10.5)
lengths <- roomDimensions[ seq(1,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES) 10 11 9 10.5
lengths

# extract just the widths into a new vector (4 5 5 5)
widths <- roomDimensions[ seq(2,length(roomDimensions),by=3) ]  # ANSWER (MULTIPLE LINES) 4 5 5 5
widths

# extract just the heights into a new vector (10 9 10.5 9.5)
heights <- roomDimensions[ seq(3,length(roomDimensions),by=3) ] # ANSWER (MULTIPLE LINES)
heights

# create a vector with the room numbers, i.e. 1 2 3 4
roomNumbers <- 1:(length(roomDimensions)/3)                     # ANSWER (MULTIPLE LINES) 1 2 3 4 
roomNumbers

# find the room numbers for the rooms that have at least 50sq feet of floor space 
desiredRooms <- roomNumbers[lengths * widths >= 50]             # ANSWER (MULTIPLE LINES)  2 3 
desiredRooms

# get the heights for just those rooms (final answer)
heights[desiredRooms]                                           # FINAL ANSWER (MULTIPLE LINES)  9 10
```
:::

---

### QUESTION 25  # TOPICS: vectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 25
# TOPICS: vectors
#
# Display the sum of the squares of the numbers in the vector nums.
# For example, if nums contains the values 4 2 and 5 then your command should
# display the number 45, (i.e. 16 + 4 + 25)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
nums <- c(4,2,5)
sum(nums^2)        # ANSWER
```
:::

---

### QUESTION 26 TOPICS: vectors, indexingWithLogicalValues, vectorAssignment


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 26
# TOPICS: vectors, indexingWithLogicalValues, vectorAssignment
#
# Grades is a vector that contains grades on a test. 
# Modify the vector, grades, by adding 5 points to all grades that are below 60.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
grades <- c(53,70,49,60,90)
grades[grades < 60] <- grades[grades < 60] + 5   # ANSWER
grades
```
:::

---

### QUESTION 27 TOPICS: vectors, indexingWithNumbers

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 27
# TOPICS: vectors, indexingWithNumbers
#
#  Reverse the contents of the vector x. DO NOT USE THE rev FUNCTION. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- seq(10,50,by=10)
x
x <- x[length(x):1]     # ANSWER
x
```
:::

---

### QUESTION 28 TOPICS: vectors, mean, length, indexingWithLogicalValues

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 28
# TOPICS: vectors, mean, length, indexingWithLogicalValues
#
# grades is a vector that contains grades on a test. Display the number of students 
# whose grade is equal to the average grade. For example, if grades contains
# c(80,85,85,85,90) then the answer should be 3 since the average is 85 and 3 students
# received exactly and 85. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
grades <- c(80,85,85,85,90)
mean(grades)
length( grades[grades == mean(grades)] )  # ANSWER
```
:::

---

### QUESTION 29 TOPICS: vectors, seqOrColon, minMax

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 29
# TOPICS: vectors, seqOrColon, minMax
#
# store in the vector nums a sequence of values from the smallest value in the
# vector x to the largest value in x. For example if x would contain c(7,8,7,3,5),
# your code should set the value of nums to the sequence 3 4 5 6 7 8
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(7,10,7,3,5)
min(x)
max(x)
nums <- min(x):max(x)   # ANSWER
nums
```
:::

---

### QUESTION 30 TOPICS: vectors, minMax, c


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 30
# TOPICS: vectors, minMax, c
#
# Write a single command that shows the lowest and the highest values from the
# vector x. For example if x contains c(20,-5,-2,40,25,-2,18), then your answer
# should display -5 40
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(20,-5,-2,40,25,-2,18)    # start with some data for x
c(min(x),max(x))                # ANSWER
```
:::

---

### QUESTION 31 TOPICS: vectors, roundTruncCeilingFloor, sqrt, ==and!= , %/%and%%


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 31
# TOPICS: vectors, roundTruncCeilingFloor, sqrt, ==and!= , %/%and%%
#
# Determine if the number stored in n is a "perfect square". For example 25 is
# a "perfect square" because 5*5 is 25. Similarly, 49 is a perfect square because
# 7*7 is 49. However, 50 is NOT a perfect square since its square root is not an integer.
# Your code should display the value TRUE is n is a perfect square and FALSE if n is
# not a perfect square. Hint: the sqrt(x) function returns the square root of x.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
n <- 25                   # let's try with a number that IS a perfect square
trunc(sqrt(n)) == sqrt(n) # ANSWER    (TRUE)
floor(sqrt(n)) == sqrt(n) # ANOTHER POSSIBLE ANSWER   (TRUE)
n%%sqrt(n) == 0           # YET ANOTHER POSSIBLE ANSWER   (TRUE)


# ------- CHECK -------
n <- 30                   # now, just to make sure, let's try with a number that is NOT a perfect square
sqrt(n)
trunc(sqrt(n))
trunc(sqrt(n)) == sqrt(n) # ANSWER   (FALSE)
floor(sqrt(n)) == sqrt(n) # ALTERNATE ANSWER   (FALSE)
n%%sqrt(n) == 0           # YET ANOTHER POSSIBLE ANSWER   (FALSE)
```
:::

---

### QUESTION 32 TOPICS: vectors, roundTruncCeilingFloor, &_|, ==_!=


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 32
# TOPICS: vectors, roundTruncCeilingFloor, &_|, ==_!=
#
# Write a command that shows the numbers in the vector nums that are both even
# and perfect squares. For example, if nums contains c(81, 6, 36, 10, 4, 25)
# then your answer should display 36  4
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x[sqrt(x)^2 == floor(x)  &  x%%2 == 0]    # ANSWER
```
:::

---

### QUESTION 33 TOPICS: vectors


```{r}
############################################################################################################.
# NOTE: The sort(x) function will return a vector with the values from the vector x in sorted order. 
#       The unique(x) function will return a copy of the vector x with all of the duplicate values removed. 
#       These functions will help you to answer the following questions
############################################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 33
# TOPICS: vectors
#
# Write a command that displays the 2nd lowest value in the vector x. If the lowest value appears
# more than once in the vector, then your code should display that value. For example, if x contains
# the numbers 7 3 3 7 3 5 20, the function should return the number 3. Even though 3 is the lowest
# value, it appears more than once, so in that sense, it is the 2nd lowest value.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(7,3,3,7,3,5,20)
sort(x)
sort(x)[2]            # ANSWER

# The following is just another example. It is NOT a new answer. 
x <- c(7,7,3,5,20) # ANOTHER EXAMPLE
sort(x)            # 
sort(x)[2]         # 
```
:::

---

### QUESTION 34 TOPICS: vectors


```{r}
############################################################################################################.
# NOTE: The sort(x) function will return a vector with the values from the vector x in sorted order. 
#       The unique(x) function will return a copy of the vector x with all of the duplicate values removed. 
#       These functions will help you to answer the following questions
############################################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 34
# TOPICS: vectors
#
# Write a command that displays the 2nd lowest value in the vector x. For this question, you
# should ignore duplicate values. For example, if x contains the numbers 7 3 3 7 3 5 20, the function
# should return the number 5. Even though 3 appears more than once, 5 is the 2nd lowest value.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# ONE ANSWER
x <- c(7,3,3,7,3,5,20)
unique(x)
sort(x)
sort(unique(x))
sort(unique(x))[2]            # ANSWER

# ANOTHER ANSWER
x <- c(7,3,3,7,3,5,20)
xWithoutLowestNumber = x[ x != min(x) ]  # ANSWER MULTIPLE LINES
min(xWithoutLowestNumber)                # ANSWER MULTIPLE LINES
```
:::

---

### QUESTION 35 TOPICS: vectors


```{r}
############################################################################################################.
# NOTE: The sort(x) function will return a vector with the values from the vector x in sorted order. 
#       The unique(x) function will return a copy of the vector x with all of the duplicate values removed. 
#       These functions will help you to answer the following questions
############################################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 35
# TOPICS: vectors
#
# Write a command that displays the n lowest values in the vector x where n is
# a variable that contains a positive integer. The values should be displayed
# in increasing order.
#
# For example if x contains c(67,2,40,5,2,99) and n contains the number 3 then
# the command should display the numbers 2 2 and 5
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
n <- 3
x <- c(67,2,40,5,2,99)
x
sort(x)
sort(x)[1:n]            # ANSWER
```
:::

---

### QUESTION 36 TOPICS: vectors


```{r}
############################################################################################################.
# NOTE: The sort(x) function will return a vector with the values from the vector x in sorted order. 
#       The unique(x) function will return a copy of the vector x with all of the duplicate values removed. 
#       These functions will help you to answer the following questions
############################################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 36
# TOPICS: vectors
#
# Same as previous question, but the numbers should be in decreasing order. 
# For example, given the values from the previous question, the answer should
# display 5 2 2. Hint, see the help for the sort command and focus on the
# "decreasing" argument to sort.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
n <- 3
x <- c(67,2,40,5,2,99)
x
sort(x)
sort(x)[1:n]
sort(sort(x)[1:n], decreasing=TRUE)   # ANSWER

# ALTERNATE ANSWER

n <- 3
x <- c(67,2,40,5,2,99)
x
x <- sort(x, decreasing=TRUE)  # ANOTHER ANSWER (MULTIPLE LINES)
x
(length(x)-n+1)
(length(x)-n+1):length(x)
x[(length(x)-n+1):length(x)]   # ANOTHER ANSWER (MULTIPLE LINES)
```
:::

---

### QUESTION 37 TOPICS: vectors


```{r}
############################################################################################################.
# NOTE: The sort(x) function will return a vector with the values from the vector x in sorted order. 
#       The unique(x) function will return a copy of the vector x with all of the duplicate values removed. 
#       These functions will help you to answer the following questions
############################################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 37
# TOPICS: vectors
#
# Same as previous question, but the numbers should be displayed in the same order
# as they appear in the original vector x. Given the same values from the previous
# question, the answer should display the numbers  2  5  2  in that order. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
n <- 3
x <- c(67,2,40,5,2,99)
sort(x)[n]
x[x <= sort(x)[n]]     # ANSWER
```
:::

---

### QUESTION 38 TOPICS: vectors


```{r}
############################################################################################.
# QUESTION_GROUP:
#
# (The next two questions are related to each other. I would have numbered them a,b,etc but 
#  I thought it would be better for each question to have its own number.)
##########################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 38
# TOPICS: vectors
#
# x and y each contain a single integer. nums is a vector that contains many numbers. 
# You are told that the value of x is guaranteed to be less than the value of y.
# Change the vector nums so that it only contains the values that are in the range
# between x and y (including x and y). For example, if x contains 3, y contains 7 and
# nums contains c(10,3,2,11,6,4,1) then your code should change the value of nums to a vector
# containing 3 6 and 4..
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# ONE POSSIBLE ANSWER
# The following answer is done in two different steps
x <- 3 # setup some example data
y <- 7 # setup some example data
nums <- c(10,3,2,11,6,4,1) # setup some example data

nums <- nums[nums>=x]      # ANSWER (MULTIPLE LINES) - nums will now contain all the values that
# are greater than or equal to x. Given the example data above, nums,
# will now contain c(10,3,11,6,4). The 2 and 1 were eliminated from nums.

nums                       # show the new values in nums
nums <- nums[nums<=y]      # ANSWER (MULTIPLE LINES) -  nums will now contain only those values 
# that are less than or equal to y. Given the example data above, nums
# will now contain c(3,6,4). The 10, and 11 were eliminated from nums.

nums                       # show the final value of nums


# ANOTHER POSSIBLE ANSWER  -  this answer is "shorter" and more "to the point". However, this answer
#                             requires the use of the & operator.
x <- 3   # setup some example data
y <- 7   # setup some example data
nums <- c(10,3,2,11,6,4,1)   # setup some example data

nums >= x      # given the example data above, this is a logical vector that contains TRUE for every value
# of nums that is >= x. Since nums contains c(10,3,2,11,6,4,1) and x contains the number 3,
# the logical vector will be c(TRUE,TRUE,FALSE,TRUE,TRUE,TRUE,FALSE)

nums <= y      # given the example data above, this is a logical vector that contains TRUE for every value
# of nums that is <= y. Since nums contains c(10,3,2,11,6,4,1) and y contains the number 7,
# the logical vector will be c(FALSE,TRUE,TRUE,FALSE,TRUE,TRUE,TRUE)
nums>=x & nums <= y   # This will combine the two logical vectors shown above using the & operator.
# That will result in a vector where there is a TRUE in the positions that had
# TRUE for both nums>=x and for num<=y (see above). In other words, the resulting
# logical vector will contain a TRUE in those positions that correspond to values
# in nums that are BOTH >=x AND <=y.
# 
#    as explained above, "nums >= x" is:    c(TRUE ,TRUE ,FALSE,TRUE ,TRUE ,TRUE ,FALSE)
#    as explained above, "nums <= y" is:    c(FALSE,TRUE ,TRUE ,FALSE,TRUE ,TRUE ,TRUE)
# 
# Therefore, "nums>=x & nums <= y" is:      c(FALSE,TRUE ,FALSE,FALSE,TRUE ,TRUE ,FALSE)

nums <- nums[nums>=x & nums <= y]  # ANSWER - "nums <-" simply means that we are going to replace 
#          the values in "nums" with the result of the code from the 
#          right hand side of the arrow.
#     
#          The code on the "Right Hand Side" (RHS) of the arrow is:
#          "nums[nums>=x & nums<=y]". The first "nums" on the RHS of 
#          the arrow indicates that we will be taking values from "nums".
#          the [brackets] that surround the rest of the code indicate 
#          that the code inside the [brackets] will be used to identify
#          which values from nums we will be "taking". As shown above
#          the code inside the brackets results in a logical vector
#          that contains a TRUE for each value that we want from nums. 
#          
#          Given the example data above, this code therefore becomes the 
#          equivalent of: 
#          nums <- nums[c(FALSE,TRUE ,FALSE,FALSE,TRUE ,TRUE ,FALSE)]
#          
#          This results in replacing nums with the 2nd 5th and 6th values
#          from nums (i.e. those positions that contains TRUE). Therefore
#          since "nums" originally contained c(10,3,2,11,6,4,1), the 
#          new value of "nums" will be c(3,6,4)
```
:::

---

### QUESTION 39 TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 39
# TOPICS: vectors
#
# Same as previous question, except for this question the value in x may be less than,
# greater than or equal to y. Hint: use the sort() function to isolate the smaller
# of x and y and the larger of x and y. Then proceed similarly to the way you answered
# the previous question. 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# ONE ANSWER:
#
# To understand the answer to this question, you must first understand the answer to the 
# previous question. Please first review the answer to the previous question before continuing.
#
# The difference between this question and the previous question is that in the previous 
# question, you were told that the value of the variable "x" is a number that is LESS THAN
# the value in the variable "y". For example, in the previous question, it would be possible
# for x to be 3 and y to be 7 or x to b 2 and y to b 5, but it would NOT be possible for x
# to be 7 and y to be 3. HOWEVER, in this version of the question, x could be ANY number
# and y could be ANY number. 
#
# Therefore when you write the code to find the values in "nums" that are between x and y 
# you know that the numbers that you desire are all greater than x and less than y. Therefore
# in the previous question, you knew that nums>=x needed to be TRUE and nums<=y needed to be TRUE.
#
# However, in this question you DON'T know which number is less, x or y. Therefore
# if x is less than y then the code from the previous question would work. For example,
# if x is 3 and y is 7 then the expression   5>=3 & 5<=7   would result in a TRUE value.
#
# However, if y were less then x (e.g. x is 7 and y is 3) then "nums>=x & nums<=y" would
# result in a FALSE value for EVERY number. For example the number 5 is obviously between
# x=7 and y=3. However, the expression    5>=7 & 5<=3 would be FALSE.
#
# One approach to answer this question is given the values in x and y, to have your code
# "figure out" which value is higher and which is lower. You can store the lower value in a
# variable named "lower" and the higher value in a variable named "higher". Then you can simply use "lower"
# value as you used "x" in the previous question and use "higher" as you used "y" in the 
# previous question. This is exactly what is done by the code below 


x <- 7   # setup some example data
y <- 3   # setup some example data
nums <- c(10,3,2,11,6,4,1)   # setup some example data

sort(c(x,y))                # "c(x,y)" is a vector that contains the values of x and y. Given the
# example data above, since x is 7 and y is 3, c(x,y) is c(7,3). 
#
# The "sort" function takes a vector and sorts its contents. So
# sort(c(7,3)) results in the vector c(3,7)

lower <- sort(c(x,y))[1]  # As described above, given the example data above, the code:
# "sort(c(x,y))" will result in the vector c(3,7). The "[1]" that
# appears after "sort(c(x,y))" extracts the first value from the 
# sorted vector. Given the data above, the first value of c(3,7) is 3.
# This number, 3, is then assigned to the variable "lower".
lower
higher<- sort(c(x,y))[2]  # In a similar way, this code assigns the 2nd value from the sorted vector
# to the variable "higher". Specifically, given the example data, the value
# 7 will be assigned to the variable "higher".
# 
# Note that if the values in x and y were initially 3 and 7 instead of 7 and 3,
# we would still get the 3 being assigned to lower and 7 being assigned to higher.
# Now that we have variables "lower" and "higher", the answer is exactly
# the same as the answer to the previous question, except that instead of using
# x and y in the answer we will use "lower" and "higher".
higher
nums <- nums[nums>=lower]       # ANSWER (MULTIPLE LINES)
nums
nums <- nums[nums<=higher]      # ANSWER (MULTIPLE LINES)
nums

# ANOTHER POSSIBLE ANSWER  -  this answer is "shorter" and more "to the point". However, this answer
#                             requires the use of the & operator. See the explanation of the answer 
#                             to the previous question for more detail.
x <- 7
y <- 3
lower <- sort(c(x,y))[1]
higher<- sort(c(x,y))[2]
nums <- c(10,3,2,11,6,4,1)
nums <- nums[nums>=lower & nums <= higher]      # ANSWER
```
:::

---

### QUESTION 40 TOPICS: vectors


```{r}
############################################################################################.
# QUESTION_GROUP:
#
# The next few questions are related to each other.
# I would have numbered them a,b,etc but I thought it would be better
# for each question to have its own number.
############################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 40
# TOPICS: vectors
#
# nums is a vector that contains several numbers. n contains a single positive integer.
# Write code that "shifts" the values in nums over to the right "n" times. For example,
# if nums contains 10 20 30 40 50 60 70 80 90 100 and n is 2 then your command should
# reset nums to 90 100 10 20 30 40 50 60 70 80. If n were 3, your command should
# reset nums to 80 90 100 10 20 30 40 50 60 70
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# The following breaks up the answer into several lines so that it is easier to understand.
# The general idea of the solution below is to figure out which values from the original vector
# will be moved to the beginning of the vector. 

nums <- seq(10,100,by=10)   # setup some example data
nums
n <- 3                      # setup some example data
positionOfLastNumber <- length(nums)-n                     #ANSWER (MULTIPLE LINES) - This is the position
#   of the value from the original vector that 
#   will be in the last position of the answer.
#   For example if nums is:
#        c(10,20,30,40,50,60,70,80,90,100)
#   and n is 3 then the final answer will be
#        c(80,90,100, 10,20,30,40,50,60,70)
#   Therefore the 70 which is the last number 
#   in the answer, is in the 7th position in nums.

positionOfLastNumber
nums[positionOfLastNumber]
firstPiece <- nums[ (positionOfLastNumber+1):length(nums)] #ANSWER (MULTIPLE LINES) - this code captures
#   the first few numbers in the answer, which
#   are the last few numbers in nums. These numbers
#   start at the position after the position we 
#   calculated above and extend until the end of
#   the nums vector. 
# 
#   For example if nums is:
#        c(10,20,30,40,50,60,70,80,90,100)
#   and n is 3 then the final answer will be
#        c(80,90,100, 10,20,30,40,50,60,70)
#   Therefore the 70 which is the last number 
#   in the answer, is in the 7th position in nums.
#   However, the 80 (which is in the 7th+1 postion,
#   (i.e. the 8th position) in nums will be in the 
#   first position in the final answer. The rest
#   of the numbers in nums from the 80 through 100
#   will be in the first few positions of the answer.
# 
#   We use the variable "firstPiece" to capture
#   those numbers that will be in the first few 
#   positions in the answer.
firstPiece
lastPiece <- nums[1:(positionOfLastNumber)]                #ANSWER (MULTIPLE LINES) -  the last few numbers
#   in the answer will start from the first positon
#   in "nums" and extend until the postion that we
#   calculated above. We store those numbers
#   in the variable, "lastPiece".
lastPiece
nums <- c(firstPiece, lastPiece)                           #ANSWER (MULTIPLE LINES) - at this point we have
#   the first few numbers for the answer stored in 
#   the variable "firstPiece" and the last few
#   numbers from the answer stored in the variable
#   "lastPiece". All that remains to be done is 
#   to create a new vector with the numbers in that
#   order and assign the result back to the variable
#   "nums".
nums          


# Alternatively, you could write the answer in one line
nums <- seq(10,100,by=10)
nums
n <- 3
nums <- c(nums[ (length(nums)-n+1):length(nums)] , nums[1:(length(nums)-n)] )   # ALTERNATIVE ANSWER
nums          
```
:::

---

### QUESTION 41 TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 41
# TOPICS: vectors
#
# Same as previous question, but the numbers at the end of nums should be
# removed and the beginning of nums should be filled in with "n" zeros. For example,
# if nums contains 10 20 30 40 50 60 70 80 90 100 and n is 2 then your command should
# reset nums to 0 0 10 20 30 40 50 60 70 80. If n were 3, your command should
# reset nums to 0 0 0 10 20 30 40 50 60 70
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
nums <- seq(10,100,by=10)
nums
n <- 3
nums <- c(rep(0,n) , nums[1:(length(nums)-n)] )  # ANSWER
nums          
```
:::

---

### QUESTION 42 TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 42
# TOPICS: vectors
#
# Same as above, but replace the numbers at the beginning of nums with 1,2,3,etc.
# For example, if nums contains 10 20 30 40 50 60 70 80 90 100 and n is 2 then your
# command should reset nums to 1 2 10 20 30 40 50 60 70 80. If n were 3, your command
# should reset nums to 1 2 3 10 20 30 40 50 60 70
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
nums <- seq(10,100,by=10)
nums
n <- 3
nums <- c(1:n , nums[1:(length(nums)-n)] )   # ANSWER
nums          
```
:::

---

### QUESTION 43 TOPICS: vectors


```{r}
############################################################################################.
# QUESTION_GROUP: 
# (The next few questions are related to each other. I would have numbered them a,b,etc but 
#  I thought it would be better for each question to have its own number.)
############################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 43
# TOPICS: vectors
#
# The vector nums contains several numbers. Display the values of nums without the
# smallest numbers. For example, if nums contains c(30,5,20,5,30,15) then your command
# should display 30 20 30 15(none of the 5's are displayed since 5 is the smallest number).
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
nums <- c(30,5,20,5,30,15) 
min(nums)
nums != min(nums)
nums[nums != min(nums)]    # ANSWER
nums[nums > min(nums)]     # ALTERNATIVE ANSWER
```
:::

---

### QUESTION 44 TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 44
# TOPICS: vectors
#
# Same as the previous question, but do not display any of the smallest or the
# largest numbers. For example, if nums contains c(30,5,20,5,30,15) then your command
# should display 20 15 (none of the 5's are displayed since 5 is the smallest number and
# none of the 30's are displayed since 30 is the largest number).
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

##############################################################################################.
# END_OF_QUESTION_GROUP
##############################################################################################.
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# The best way to do this is with the "&" operator. If you haven't learned the "&" operator yet
# There are alternatives below
nums <- c(30,5,20,5,30,15) 
nums[nums != min(nums) & nums != max(nums)]   # ANSWER
nums[nums > min(nums) & nums < max(nums)]     # ALTERNATIVE ANSWER

# This version of the answer does NOT use the & operator, but requires two lines of code.
# These alternatives use the != (i.e. "not equal to") operator
nums <- c(30,5,20,5,30,15) 
temp <- nums[nums != min(nums)]  #ANSWER (MULTIPLE LINES)
temp[temp != max(nums)]          #ANSWER (MULTIPLE LINES)

# This version of the answer does NOT use the & operator, but requires two lines of code.
# These alternatives use the > and < operators instead of the != operator
nums <- c(30,5,20,5,30,15) 
temp <- nums[nums > min(nums)]  #ALTERNATIVE ANSWER (MULTIPLE LINES)
temp[temp < max(nums)]          #ALTERNATIVE ANSWER (MULTIPLE LINES)
```
:::

---

### QUESTION 45 TOPICS: vectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 45
# TOPICS: vectors
#
# x and y are two vectors whose lengths are the same. Merge the contents of
# vector y into vector x in the following way. If x1,x2,x3,x4,... represent the
# initial values in x and y1,y2,y3,y4,... represent the values in y, then after
# your commands run, the values in x should be x1,y1,x2,y2,x3,y3,x4,y4,...
#
# For example, if x contains the values 1 2 3 4 and y contains the values 11 22 33 44 
# then after your code runs the values in x should be 1 11 2 22 3 33 4 44. 
# (You may use more than one command to accomplish this result).
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x <- c(1,2,3,4)
x
y <- c(11,22,33,44)
y
seq(1,2*length(x),by=2)


x[seq(1,2*length(x),by=2)] <- x    # ANSWER (MULTIPLE LINES)

x
seq(2,2*length(y),by=2)


x[seq(2,2*length(y),by=2)] <- y    # ANSWER (MULTIPLE LINES)
x
```
:::

## Dataframes


```{r}
##############################################################################################.
##############################################################################################.
##
##            TOPIC:  DATAFRAMES
##
##############################################################################################.
##############################################################################################.


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 46
# TOPICS: dataframes
#
# Write R code that creates a dataframe that has 2 columns.
#
# The first column should contain the even numbers 
# from 2 through 1000, ie. 2,4,6 ... 1000.
#
# The 2nd column should contain the odd numbers
# from 999 counting down to 1, i.e. 999, 997, 995, ... 1
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
##############.
# ANSWER 1
##############.

# This answer creates the data.frame in one command.
# NOTE: strinsAsFactors=FALSE is not necessary here since there are no character vectors in this dataframe.
numbers <- data.frame(evens=seq(2,1000,by=2), odds = seq(999,1,by=-2))   # ANSWER

# show the first 3 and last 3 rows of numbers (this is not part of the answer)
head(numbers,3)   # show the first 3 rows
tail(numbers,3)   # show the last 3 rows

##############.
# ANSWER 2 
##############.

# This answer creates variables for the vectors first.
# The column names will be the same as the variable names of the vectors.
evens <- seq(2,1000,by=2)              # ANSWER (MULTIPLE LINES)
odds <- seq(999,1, by=-2)              # ANSWER (MULTIPLE LINES)
numbers <- data.frame(evens,odds)      # ANSWER (MULTIPLE LINES)

# show the first 3 and last 3 rows of numbers (this is not part of the answer)
head(numbers,3)   # show the first 3 rows
tail(numbers,3)   # show the last 3 rows
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 47
# TOPICS: dataframes
#
# Write R code that creates a dataframe that contains the following information:
#  
# The name of the 1st column should be: "department"
# The values in this column should be: produce, bakery, produce, produce, bakery. 
#
# The name of the 2nd column should be: "product"
# The values in this column should be: apple, chocolate cake, orange, pear, rye bread. 
#
# The name of the 3rd column should be: "price"
# The values in this column should be: 1.99 , 1.59, 0.99, 1.99, 4.99.
#
# Store the dataframe in a variable named "prices".
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
###########################################################.
# ANSWER 1 - create the data.frame with a single command
###########################################################.

prices <- data.frame(department=c("produce","bakery","produce","produce","bakery"),   # ANSWER (MULTIPLE LINES)
                     product=c("apple","chocolate cake","orange","pear","rye bread"), # ANSWER (MULTIPLE LINES)
                     price=c(1.99 , 1.59, 0.99, 1.99, 4.99),                          # ANSWER (MULTIPLE LINES)
                     stringsAsFactors=FALSE)                                          # ANSWER (MULTIPLE LINES)

# check our answer (this is NOT part of the answer)
prices

###########################################################.
# ANSWER 2 - create variables for the columns first
###########################################################.

department <- c("produce","bakery","produce","produce","bakery")        # ANSWER (MULTIPLE LINES)
product <- c("apple","chocolate cake","orange","pear","rye bread")      # ANSWER (MULTIPLE LINES)
price <- c(1.99 , 1.59, 0.99, 1.99, 4.99)                               # ANSWER (MULTIPLE LINES)
prices <- data.frame(department,product,price,stringsAsFactors = FALSE) # ANSWER (MULTIPLE LINES)

# check our answer (this is NOT part of the answer)
prices
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 48
# TOPICS: dataframes
#
# Write a command that changes the names of the columns in the dataframe
# from the previous question. The first column should now be named itemName and
# the 2nd column should be named pricePerLb. Hint: use the colnames function.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
###############.
# ANSWER 1
###############.
colnames(prices)[c(2,3)] <- c("itemName","pricePerLb")  # ANSWER
# check ...
prices

################.
# ANSWER 2
################.
colnames(prices)[2] <- "itemName"     # ANSWER (MULTIPLE LINES)
colnames(prices)[3] <- "pricePerLb"   # ANSWER (MULTIPLE LINES)
# check ...
prices
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 49
# TOPICS: dataframes
#
# Challenge: Write a command that changes the names of the rows in the
# "prices" dataframe from the previous question. The new names of the rows
# should be item1, item2 , etc. Write the code so that no matter how many rows
# the dataframe would contain the names of the rows would follow the same pattern.
# For example, if the dataframe happened to contain 1000 rows, then after your
# command executed, they would be named: item1,item2,item3,item4,... item1000
#
# Hint: As part of your answer, you will need to use the paste function
# with sep="" (nothing between the quotes). Note that we did not cover the paste
# function in class. You should look at the help for the paste function and try a
# few examples of your own. You can also type the R command example(paste) to
# automatically run the examples that appear at the end of the help page for the
# paste function.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the prices dataframe.

#############################.
# sample answer:
#############################.

# To generate the new rownames you can use the following command:
paste("item", 1:length(rownames(prices)), sep="")

# or the following command
paste("item", 1:nrow(prices), sep="")

# now just assign the result of one of these commands to the rownames:
rownames(prices) <- paste("item", 1:nrow(prices), sep="")    # ANSWER  (THIS IS THE ONLY LINE YOU NEED)

# check:
prices
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 50 (a)
# TOPICS: dataframes
#
# Write a command that modifies the dataframe from the previous question.
# The command should change the word "apple" to "mcintosh apple".
#
# a. Do this assuming that the dataframe was created with the option stringsAsFactors=FALSE.
#     For this simple answer, you may refer to the row and column number if you like.
#
# b. Do this again, but this time DO NOT refer to the actual row number or column number in your answer. Your answer should work successfully, no matter what row contains "apple". 
#     (NOTE: This is an important technique that will be used later in the course to create our own functions).
#
# c. Do this assuming that the dataframe was created with the option stringsAsFactors=TRUE.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

# ANSWER - part a
prices[1,2] <- "mcintosh apple"    # ANSWER

# ANSWER - part b
prices[prices$itemName == "apple" , "itemName"] <- "mcintosh apple"    # ANSWER
prices

# ANSWER - part c

################.
# ANSWER 1
################.

# Note that based on the above commands, the first two columns in the dataframe
# are FACTORS. We can see the "levels" of the 2nd column with the following command:
levels(prices[[2]])
# since "mcintosh apple" is NOT a valid level in the 2nd column, the following
# command will generate an error and will replace the word "apple" NOT with 
# "mcintosh apple", but rather with "NA"
#
#      prices[1,2] <- "mcintosh apple"
#
# To avoid this problem, we can add "mcintosh apple" to the levels of the 2nd
# columnn and only then change the value of "apple" to "mcintosh apple".
levels(prices[[2]]) <- c(levels(prices[[2]]),"mcintosh apple")   # ANSWER (MULTIPLE LINES)
levels(prices[[2]])
prices[1,2] <- "mcintosh apple"                                  # ANSWER (MULTIPLE LINES)
#check the answer
prices

################.
# ANSWER 2  (read ANSWER 1 first)
################.

# Another way to avoid the problem is to convert the 2nd column from a factor
# into a vector of characters 
prices[[2]] <- as.character(prices[[2]])      # ANSWER (MULTIPLE LINES)
# check - this will now be a vector
prices[[2]]
# now we can change "apple" to "mcintosh apple" without worrying about levels
prices[1,2] <- "mcintosh apple"               # ANSWER (MULTIPLE LINES)
prices
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 51
# TOPICS: dataframes
#
# Write a command that modifies the dataframe from the previous question.
# The command should increase the pricePerLb of oranges by 10%. Round the increased
# price to the nearest penny.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the prices dataframe.

###############################.
# ANSWER 1 
#    - using $ notation  
#    - answer the question using two commands (may be easier to understand than a single command)
###############################.
# first let's identify which row contains oranges:
rows <- prices$itemName == "orange"                                      # ANSWER (MULTIPLE LINES)
rows

# now lets change the price for that row. Note that we are rounding the new price to the nearest penny.
prices[rows,"pricePerLb"] <- round( prices[rows,"pricePerLb"] * 1.1, 2)  # ANSWER (MULTIPLE LINES)

# check our answer
prices


#########################################.
# ANSWER 2
#    - using [[double-bracket]] notation
#    - answer the question using two commands (may be easier to understand than a single command)
#########################################.
# first let's identify which row contains oranges:
rows <- prices[["itemName"]] == "orange"                                 # ANSWER (MULTIPLE LINES)
rows

# now lets change the price for that row. Note that we are rounding the new price to the nearest penny.
prices[rows,"pricePerLb"] <- round( prices[rows,"pricePerLb"] * 1.1, 2)  # ANSWER (MULTIPLE LINES)

# check our answer
prices



#########################################.
# ANSWER 3
#    - using $ notation  
#    - answer using ONE command (avoids the need to create new variables that are otherwise unnecessary)
#########################################.

# we can avoid the need to create a new variable using the following approach to 
# answer the question using a single command (for readability, the command is typed on two lines)
prices[prices$itemName=="orange","pricePerLb"] <-                    # ANSWER (MULTIPLE LINES)
  round(prices[prices$itemName=="orange","pricePerLb"]*1.1, 2) # ANSWER (MULTIPLE LINES)

# check our answer
prices
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 52
# TOPICS: dataframes
#
# Write a command that modifies the "prices" dataframe from the previous question.
# The command should increase the price of all items by 10%. Round the increased prices
# to the nearest penny.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the prices dataframe.

##############################.
# ANSWER 1 - using $ notation
##############################.

prices$pricePerLb <- round( prices$pricePerLb * 1.1 , 2 )              # ANSWER

# check
prices

###############################################.
# ANSWER 2 - using [[double-bracket]] notation
###############################################.
prices[["pricePerLb"]] <- round ( prices[["pricePerLb"]] * 1.1 , 2)   # ANSWER

# check
prices
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 53
# TOPICS: dataframes
#
# Refer to the dataframe from the previous question. Write a command that adds
# a new column named salePrice. The new column should contain the values from
# the price column reduced by 25%. Round the prices in the new column to the nearest penny.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the prices dataframe.

###############################################.
# ANSWER 1 - using $ notation
###############################################.

prices$salePrice <- round ( prices$pricePerLb * 0.75 , 2)     # ANSWER

# check 
prices

###############################################.
# ANSWER 2 - using [[double-bracket]] notation
###############################################.

prices[["salePrice"]] <- round ( prices[["pricePerLb"]] * 0.75 , 2)   # ANSWER

# check 
prices
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 54
# TOPICS: dataframes
#
# Refer to the dataframe from the previous question.
# Write a command that displays the itemName and price of all items from the
# produce department. Write the code so that it will continue work even if the
# specific data in the dataframe changes.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the prices dataframe.

###############################################.
# ANSWER 1 - using $ notation
###############################################.

# First identify the rows we want:
rows <- prices$department == "produce"
rows

# now get the data for those rows from the specified columns:
prices[rows,c("itemName","pricePerLb")]
#
# Alternatively, we can do all this in one line:
prices[     prices$department == "produce"    ,c("itemName","pricePerLb")]

###############################################.
# ANSWER 2 - using [[double-bracket]] notation
###############################################.

# First identify the rows we want:
rows <- prices[["department"]] == "produce"
rows

# now get the data for those rows from the specified columns:
prices[rows,c("itemName","pricePerLb")]

# Alternatively, we can do all this in one line:
prices[     prices[["department"]] == "produce"    ,c("itemName","pricePerLb")]
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 55
# TOPICS: dataframes
#
# Refer to the dataframe from the previous question.
# Write a command that displays the itemName and price of all items whose price
# is 2.00 or more. Write the code so that it will continue work even if the
# specific data in the dataframe changes.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

###############################################.
# ANSWER 1 - using $ notation
###############################################.

# First identify the rows we want:
rows <- prices$pricePerLb >= 2
rows

# now get the data for those rows from the specified columns:
prices[rows,c("itemName","pricePerLb")]

# Alternatively, we can do all this in one line:
prices[     prices$pricePerLb >= 2    ,c("itemName","pricePerLb")]

###############################################.
# ANSWER 2 - using [[double-bracket]] notation
###############################################.

# First identify the rows we want:
rows <- prices[["pricePerLb"]] >= 2
rows
# now get the data for those rows from the specified columns:
prices[rows,c("itemName","pricePerLb")]
#
# Alternatively, we can do all this in one line:
prices[     prices[["pricePerLb"]] >= 2    ,c("itemName","pricePerLb")]
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 56
# TOPICS: dataframes
#
# Refer to the dataframe from the previous question.
# Write a command that displays all of the columns for all items whose
# price is 2.00 or more. Write the code so that it will continue work
# even if the specific data in the dataframe changes.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

###############################################.
# ANSWER 1 - using $ notation
###############################################.

# First identify the rows we want:
rows <- prices$pricePerLb >= 2
rows

# now get the data for those rows from the specified columns:
prices[rows,]

# Alternatively, we can do all this in one line:
prices[     prices$pricePerLb >= 2  ,   ]

###############################################.
# ANSWER 2 - using [[double-bracket]] notation
###############################################.

# First identify the rows we want:
rows <- prices[["pricePerLb"]] >= 2
rows

# now get the data for those rows from the specified columns:
prices[rows,]

# Alternatively, we can do all this in one line:
prices[     prices[["pricePerLb"]] >= 2  ,   ]
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 57
# TOPICS: dataframes
#
# Refer to the dataframe from the previous question. 
# Write a command that calculates the average pricePerLb of all
# items in the produce department.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

###############################################.
# ANSWER 1 - using $ notation
###############################################.
# First identify the rows we want:
rows <- prices$department == "produce"
rows

# now get the data for those rows from the specified columns:
mean( prices[rows,"pricePerLb"] )

# Alternatively, we can do all this in one line:
mean( prices[   prices$department == "produce"   ,"pricePerLb"] )

###############################################.
# ANSWER 2 - using [[double-bracket]] notation
###############################################.
# First identify the rows we want:
rows <- prices[["department"]] == "produce"
rows

# now get the data for those rows from the specified columns:
mean( prices[rows,"pricePerLb"] )

# Alternatively, we can do all this in one line:
mean( prices[   prices[["department"]] == "produce"   ,"pricePerLb"] )
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 58
# TOPICS: dataframes
#
# Refer to the dataframe from the previous question. 
# Write a command that displays all of the data from just the odd numbered rows.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

##########.
# ANSWER 
##########.

# First identify the rows we want:
rows <- seq(1,nrow(prices),by=2)
rows

# now get the data for those rows from the specified columns:
prices[ rows , ]

# Alternatively, we can do all this in one line:
prices[  seq(1,nrow(prices),by=2)   ,  ]
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 59
# TOPICS: dataframes
#
# Refer to the dataframe from the previous question.
# Write a command that displays just the name of the items whose
# pricePerLb is at least 1.00 but not more than 2.00.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

###############################################.
# ANSWER 1 - using $ notation
###############################################.

# First identify the rows we want:
rows <- prices$pricePerLb >= 1 & prices$pricePerLb <= 2
rows

# now get the data for those rows from the specified columns:
prices[ rows , "itemName"]

# Alternatively, we can do all this in one line:
prices[    prices$pricePerLb >= 1 & prices$pricePerLb <= 2    , "itemName"]

###############################################.
# ANSWER 2 - using [[double-bracket]] notation
###############################################.
# First identify the rows we want:
rows <- prices[["pricePerLb"]] >= 1 & prices[["pricePerLb"]] <= 2
rows

# now get the data for those rows from the specified columns:
prices[ rows , "itemName"]

# Alternatively, we can do all this in one line:
prices[    prices[["pricePerLb"]] >= 1 & prices[["pricePerLb"]] <= 2    , "itemName"]
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 60
# TOPICS: dataframes
#
# Refer to the dataframe from the previous question.
# Write a command that displays just the name of the items whose
# pricePerLb is either less than 1.00 or more than 2.00.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

###############################################.
# ANSWER 1 - using $ notation
###############################################.

# First identify the rows we want:
rows <- prices$pricePerLb < 1 | prices$pricePerLb > 2
rows

# now get the data for those rows from the specified columns:
prices[ rows , "itemName"]

# Alternatively, we can do all this in one line:
prices[    prices$pricePerLb < 1 | prices$pricePerLb > 2    , "itemName"]

###############################################.
# ANSWER 2 - using [[double-bracket]] notation
###############################################.

# First identify the rows we want:
rows <- prices[["pricePerLb"]] < 1 | prices[["pricePerLb"]] > 2
rows

# now get the data for those rows from the specified columns:
prices[ rows , "itemName"]

# Alternatively, we can do all this in one line:
prices[    prices[["pricePerLb"]] < 1 | prices[["pricePerLb"]] > 2    , "itemName"]
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 61
# TOPICS: dataframes
#
# Refer to the dataframe from the previous question.
# Write a command that displays just the name of the produce whose pricePerLb is at
# least 1.00 but not more than 2.00.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

###############################################.
# ANSWER 1 - using $ notation
###############################################.

# First identify the rows we want:
rows <- prices$department == "produce" & prices$pricePerLb >= 1 & prices$pricePerLb <= 2
rows

# now get the data for those rows from the specified columns:
prices[ rows , "itemName"]

# Alternatively, we can do all this in one line:
prices[prices$department == "produce" & prices$pricePerLb >= 1 & prices$pricePerLb <= 2      , "itemName"]


###############################################.
# ANSWER 2 - using [[double-bracket]] notation
###############################################.

# First identify the rows we want:
rows <- prices[["department"]] == "produce" & prices[["pricePerLb"]] >= 1 & prices[["pricePerLb"]] <= 2
rows

# now get the data for those rows from the specified columns:
prices[ rows , "itemName"]

# Alternatively, we can do all this in one line:
prices[prices[["department"]]=="produce" & prices[["pricePerLb"]]>=1 & prices[["pricePerLb"]]<=2  , "itemName"]
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 62
# TOPICS: dataframes
#
# Refer to the dataframe from the previous question.
# Write a single command that displays 
#      the name of the produce whose pricePerLb is less than 1.00 and also 
#      the produce whose pricePerLb is greater than 2.00
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

###############################################.
# ANSWER 1 - using $ notation
###############################################.
# First identify the rows we want (notice the parentheses - they ARE necessary)
rows <- prices$department == "produce" & ( prices$pricePerLb < 1 | prices$pricePerLb > 2 )
rows

# now get the data for those rows from the specified columns:
prices[ rows , "itemName"]

# Alternatively, we can do all this in one line:
prices[prices$department=="produce" & (prices$pricePerLb<1 | prices$pricePerLb>2 ) , "itemName"]

###############################################.
# ANSWER 2 - using [[double-bracket]] notation
###############################################.
# First identify the rows we want (notice the parentheses - they ARE necessary)
rows<-prices[["department"]]=="produce" & (prices[["pricePerLb"]]<1 | prices[["pricePerLb"]]>2)
rows

# now get the data for those rows from the specified columns:
prices[ rows , "itemName"]

# Alternatively, we can do all this in one line:
prices[prices[["department"]]=="produce" & (prices[["pricePerLb"]]<1 | prices[["pricePerLb"]]>2 ) , "itemName"]
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 63
# TOPICS: dataframes
#
# Write a command that displays all of the data for a "randomly chosen" row
# from the dataframe. Your answer should work even if the data in the dataframe changes.
#
# Hint: as part of your answer, use the nrow function to determine the number of rows
# that are in the dataframe.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

###############.
# Answer:
###############.

# First identify a random row number (between 1 and the one more than the
# max # of rows in the dataframe). The "+1" is because we will be "truncating" this
# number in the next step.
row <- runif(1,1,nrow(prices) + 1)
row

# now "truncate" that result to get an integer
row <- trunc(row)
row

# now display the data for the randomly chosen row:
prices[ row , ]

# Alternatively, we can do all this in one line:
prices[   trunc(runif(1,1,nrow(prices) + 1))   ,   ]
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 64
# TOPICS: dataframes
#
# Write a command that displays the average pricePerLb of those items
# whose pricePerLb is at least 1.50
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

###############################################.
# ANSWER 1 - using $ notation
###############################################.

# First identify the rows we want
rows <- prices$pricePerLb >= 1.5
rows

# now get the data for those rows from the specified columns:
mean( prices[ rows , "pricePerLb"] )

# Alternatively, we can do all this in one line:
mean( prices[ prices$pricePerLb >= 1.5 , "pricePerLb"] )

###############################################.
# ANSWER 2 - using [[double-bracket]] notation
###############################################.
# First identify the rows we want
rows <- prices[["pricePerLb"]] >= 1.5
rows

# now get the data for those rows from the specified columns:
mean( prices[ rows , "pricePerLb"] )

# Alternatively, we can do all this in one line:
mean( prices[ prices[["pricePerLb"]] >= 1.5 , "pricePerLb"] )
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 65
# TOPICS: dataframes
#
# Write a command that displays the number of items that are listed in
# the "produce" department.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

#########.
# ANSWER 
#########.

# First identify the rows we want as a logical vector
rows <- prices$department == "produce"
rows

# Now sum the logical vector to get the total number of TRUEs
sum(rows)

# Alternatively, we can do all this in one line:
sum(  prices$department == "produce"  )
```
:::

---

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 66
# TOPICS: dataframes
#
# Write a command that displays the number of items that are listed
# in the "produce" department whose pricePerLb is at least 1.50.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# NOTE: see the answers to questions 47 and 48 for code to create the dataframe.

#########.
# ANSWER 
#########.

# First identify the rows we want as a logical vector
rows <- prices$department == "produce"  &  prices$pricePerLb >= 1.5
rows

# Now sum the logical vector to get the total number of TRUEs
sum(rows)

# Alternatively, we can do all this in one line:
sum(     prices$department == "produce"  &  prices$pricePerLb >= 1.5    )
```
:::

## User Defined Functions

### QUESTION 67 TOPICS: user defined functions


```{r}
##############################################################.
##############################################################.
##
##   TOPIC: USER DEFINED FUNCTIONS  (i.e. creating your own functions
##     For each of the following questions, write a function that
##     matches the description in the comment and that 
##     has the "signature" shown below the comment.
##
##############################################################.
##############################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 67
# TOPICS: user defined functions
#
# getGpa calculates a student's Grade Point Average (GPA)
# as a "weighted average" of the grades for the courses
# the student has taken, weighted by the number of credits
# for each course. 
#
# The grades argument is expected to be a vector that 
# contains one numeric grade for each course. The expected 
# numeric values are 4,3,2,1 or 0 for grades of A,B,C,D or F 
# respectively. 
# 
# The credits argument is expected to be a vector the same 
# length as the grades argument. The values in the credits
# vector represent the credits for each course from the 
# corresponding grades vector.
# 
# The function should return a single number that represents
# the GPA.
#
# Example: The following function call is used to calculate
# the GPA of a student who took 3 classes. The grades in the
# three classes were A,B,B (i.e. 4,3,3) and the credits for
# the classes were 2, 4 and 3 credits respectively.
# This is calculated as (4*2 + 3*4 + 3*3)/9  to get 3.222222
#
#    > getGpa(c(4,3,3), c(2,4,3))
#    3.222222
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
getGpa <- function(grades , credits) {
  sum(grades * credits) / sum(credits) 
}
```
:::

---

### QUESTION 68a. TOPICS: user defined functions


```{r}
############################################################################################.
# QUESTION_GROUP: 68
# TOPICS: user defined functions
#
# The next few questions are related to each other.
#
# For each part below, create the function as described
############################################################################################.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 68a.
# TOPICS: user defined functions
#
# ARGUMENTS/PARAMETERS 
#
#   vec - a vector that contains 2 or more values
#
# Returns a vector that contains the two largest
# values from the vector,vec . The values should 
# be returned IN INCREASING ORDER regardless
# of the order in the original vector. For example:
#     > getTopTwoValues(c(1,4,2,5,3))
#     4  5
#     > getTopTwoValues(c(1,5,2,4,3))
#     4  5
# If the largest value in vec appears more than once
# then the return value should include that value twice.
# For example:
#     > getTopTwoValues(c(1,4,2,4,3,4))
#     4  4
#
# Hints: You can write the function in many different ways. 
# Some functions that might help are: sort, head, tail 
#
# Signature:    getTopTwoValues <- function(vec) 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
getTopTwoValues <- function(vec) {
  sorted <- sort(vec)
  sorted[ c ( length(vec)-1 , length(vec) ) ]
}
```
:::

---

### QUESTION 68b. TOPICS: user defined functions


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 68b.
# TOPICS: user defined functions
#
# Same as previous question but this function takes a parameter
# named "n" that indicates how many values to return. "n" has a default
# value of 2. Also vec, may contain 1 or more values instead of 2 or more values.
# If there are fewer values in vec than "n" then the function should
# return all of the values from vec.
#
# ARGUMENTS/PARAMETERS 
#
#   vec - any vector
#
#   n - the number of largest values to return
#
# Returns a vector that contains the "n" largest
# values from the vector. If "n" is larger than the number of 
# values in vec then the function should return all of the 
# values in vec.
#
# Examples:
#     > getTopValues(c(1,9,2,5,3),3)
#     3  5  9
#     > getTopValues(c(1,5,2,9,3))
#     5  9
#     > getTopValues(c(1,99,55,2,99),3)
#     55  99  99
#     > getTopValues(c(1,99),3)
#     1  99
#
# SIGNATURE:  getTopValues <- function(vec, n=2)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# ONE ANSWER
getTopValues <- function(vec, n=2) {
  n = min( n , length(vec) )  # ensure n is not longer than length(vec)
  sorted <- sort(vec)
  position <- length(vec)-(n-1)
  sorted[ position : length(vec) ]
}

# Test the answer
getTopValues(c(1,9,2,5,3),3)   #     3  5  9
getTopValues(c(1,5,2,9,3))     #     5  9
getTopValues(c(1,99,55,2,99),3)#     55  99  99
getTopValues(c(1,99),3)        #     1  99



# ANOTHER ANSWER
getTopValues <- function(vec, n=2) {
  sorted <- sort(vec)
  position <- length(vec)-(n-1)
  if (position < 1) {
    position <- 1
  }
  sorted[ position : length(vec) ]
}

# Test the answer
getTopValues(c(1,9,2,5,3),3)   #     3  5  9
getTopValues(c(1,5,2,9,3))     #     5  9
getTopValues(c(1,99,55,2,99),3)#     55  99  99
getTopValues(c(1,99),3)        #     1  99
```
:::

---

### QUESTION 68c TOPICS: user defined functions


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 68c
# TOPICS: user defined functions
#
# same as previous question except the values that are 
# returned should NOT be the same, unless, the vector contains 
# too few unique values
#
# For example:
#     > getTopValues(c(1,9,3,9), 2)
#     3  9
#     > getTopValues(c(9,9,9,9), 2)
#     9  9
#     > getTopValues(c(9,1,9,2), 3)
#     1  2  9
#
# Hints: You can write the function in many different ways. 
# The following are some functions that might help. If you are
# not familiar with any of these functions, search the R help
# pages: unique, sort, head, tail
#
# FUNCTION SIGNATURE: getTopValues <- function(vec, n=2)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
getTopValues <- function(vec, n=2) {
  sortedUnique <- sort(unique(vec))
  numValues <- min(n,length(sortedUnique))
  sortedUnique[ (length(sortedUnique)-(numValues-1)) : length(sortedUnique) ]
}
```
:::

---

### QUESTION 68d TOPICS: user defined functions


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 68d
# TOPICS: user defined functions
#
# The following is similar to part (a) above, 
# with the following change: "The largest values should
# appear IN THE SAME ORDER AS THOSE VALUES APPEAR IN vec."
# See the description below for more info:
#
# ARGUMENTS/PARAMETERS 
#
#   vec - a vector that contains 2 or more values
#
# Returns a vector that contains the two largest
# values from the vector,vec . The largest values
# should appear IN THE SAME ORDER AS THOSE VALUES
# APPEAR IN vec. For example:
#     > getTopTwoValues(c(1,9,2,5,3))
#     9  5
#     > getTopTwoValues(c(1,5,2,9,3))
#     5  9
# If the largest value in vec appears more than once
# then the return value should include that value twice.
# For example:
#     > getTopTwoValues(c(1,9,2,9,3,9,9))
#     9  9
#
# FUNCTION SIGNATURE: getTopTwoValues <- function(vec)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
################.
#  68d - ANSWER
################.

# The following solution works in the following way to make sure
# that the 2 values that are returned are in the same order
# as in the original list:
#
# "Psuedo-code" for the answer:
#
#     First find the POSITION number of the highest value. 
#     
#     Copy the vector,x , into another variable, y.
#
#     In the new copy of the vector, REPLACE the highest
#     value with the min of the vector.
#
#     Then get the POSITION number of the next highest value
#     in the same way.
#
#     Now you know the position numbers of the two highest
#     values. You should be able to get just the values from
#     those two positions in the original vector.

getTopTwoValues <- function(x) {
  positions <- 1:length(x)
  positionOfTopValue <- positions[ x == max(x) ][1]
  y <- x       # save a copy of x in y
  y[positionOfTopValue] <- min(x)
  positionOf2ndValue <- positions[ y ==  max(y) ][1]
  x[positions==positionOfTopValue | positions==positionOf2ndValue]
}

##############################################################################################.
# END_OF_QUESTION_GROUP: 68
##############################################################################################.
```
:::

---

### QUESTION 69. TOPICS: user defined functions, if_elseif_else


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 69.
# TOPICS: user defined functions, if_elseif_else
#
# For the purposes of the following questions, the following definition of
# "prime number" is assumed:   "An integer greater than one is called a prime number
# if its only positive divisors (factors) are one and itself."  Therefore, 1 is not
# a prime and neither is any negative number.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 69a.
#
# ARGUMENTS/PARAMETERS:
# 
#    num - a single numerical value
#
# Returns TRUE if num is a prime number and returns FALSE if num is not a prime number 
#
# Examples:
#    > is.prime1(-1)
#    FALSE 
#    > is.prime1(1)
#    FALSE 
#    > is.prime1(2)
#    TRUE 
#    > is.prime1(7)
#    TRUE
#    > is.prime1(9)
#    FALSE 
#
# FUNCTION SIGNATURE:   is.prime1 <- function( num )
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
is.prime1 <- function( num ) {
  if (num < 2) {
    return (FALSE)
  } else if(num == 2) {
    return (TRUE)  # the code in the else will not work for 2 (try it to see why)
  } else {
    divisors <- 2:(num/2)    # note, the following is more efficient: 2:sqrt(num)    
    remainders <- num %% divisors
    numberOfRemaindersThatAreZero <- sum (remainders == 0)
    return (numberOfRemaindersThatAreZero == 0)
  }
}
```
:::

---

### QUESTION 69b. TOPICS: user defined functions, sapply_lapply

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 69b.
# TOPICS: user defined functions, sapply_lapply
#
# ARGUMENTS/PARAMETERS:
# 
#    vec - a numeric vector 
#
# Returns a logical vector that describes if each corresponding
# number in vec is a prime number. 
#
# Examples:
#    > is.prime(c(1,2,3,4,5,6,7))
#    FALSE  TRUE  TRUE  FALSE  TRUE  FALSE  TRUE
#
#    > is.prime(c(107,109,111,1000))   # note 111 is divisble by 3 (3*37 is 111)
#    TRUE  TRUE  FALSE  FALSE
#
# HINT: You can actually use the sapply and the lapply functions with a vector 
# (technically a "vector" is known as an "atomic list"). Use the sapply function 
# and the is.prime1 function from the previous question to generate the answer
# for this function. (Why did I say to use the sapply and not the lapply function???)
#
# FUNCTION SIGNATURE:   is.prime <- function( vec )
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
is.prime <- function(vec) {
  sapply(vec,is.prime1)
}
```
:::

---

### QUESTION 69c. TOPICS: user defined functions, loops


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 69c.
# TOPICS: user defined functions, loops
#
# ARGUMENTS/PARAMETERS:
#
#    n - return all primes that are less than or equal to this value
#
# Returns a vector that contains all of the prime numbers that are less than or
# equal to "n".
# 
# EXAMPLES:
#    > primes.up.to(15)
#    2  3  5  7  11  13
#
# FUNCTION SIGNATURE:   primes.up.to <- function(n = 100)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
#################################################################.
# ANSWER #1  -  BEST ANSWER
# The following version is the best. It does not use any loops.
# However, you should still understand the code for the other
# answers below.
#################################################################.
primes.up.to <- function(n = 100){
  tfVector <- is.prime(2:n)   # tfVector will be a vector of TRUE/FALSE values
  return ( (2:n)[tfVector] )  # Return the numbers whose values from tfVector are TRUE. NOTE: the word "return" is optional.
}

#################################################################.
# ANSWER #2 
# The following version uses a "for" loop
#################################################################.
primes.up.to <- function(n = 100){
  primes <- vector()          # start with an empty vector
  
  for (num in 2:n) {          # check each number from 2 until n
    if (is.prime1(num)) {     # if the current value of num is a prime number ...
      primes <- c(primes,num)   # ... add it to the vector of primes
    }
  }
  
  return (primes)        # the word "return" is optional. The following will work too:   primes  
}


#################################################################.
# ANSWER #3
# The following version uses a "while" loop
#################################################################.
primes.up.to <- function(n = 100){
  primes <- vector()          # start with an empty vector
  num <- 2                    # set num to the first value to check
  
  while  (num <= n) {          # keep checking as long as num is less than n
    if (is.prime1(num)) {      # if the current value of num is a prime number ...
      primes <- c(primes,num)  # ... add it to the vector of primes
    }
    
    num <- num + 1  # get the next num to check. Note that this
    # must be done OUTSIDE of the if since this should be done
    # whether or not the last number checked turned out to be prime or not.
  }
  
  return (primes)        # the word "return" is optional. The following will work too:   primes  
}
```
:::

---

### QUESTION 69d. TOPICS: user defined functions, loops

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 69d.
# TOPICS: user defined functions, loops
#
# ARGUMENTS/PARAMETERS:
#
#    n - the number of prime numbers to return
#
# Returns a vector that contains the first "n" prime numbers>
#
# FUNCTION SIGNATURE:   first.n.primes <- function(n = 10)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
get.first.n.primes <- function(n) {
  
  # The next few lines prepare the variables that will be needed by the while loop below
  
  answer <- vector()      # setup an empty vector to contain the primes
  nextNumberToCheck <- 2  # The first number we will check to see if it is a prime is 2. 
  # (Yes I know that 2 is a prime, but you have to start somewhere ...)
  numberPrimesFoundSoFar <- 0  # Before we start the loop, we havent "found" any primes yet.
  
  while(numberPrimesFoundSoFar < n) {  # If we haven't "found" n primes yet, keep looping to find another prime.
    if (is.prime1(nextNumberToCheck)){  # check to see if the "nextNumberToCheck" is a prime number
      answer <- c(answer,nextNumberToCheck) # If the number is a prime then add it to the "answer" vector.
      # Note that the line "answer <- vector()" that appears above was
      # necessary so that this line works the first time it is run. 
      # If "answer" wasn't set to something at the top of this function 
      # then the first time we would execute this line we would get
      # an error since "answer" in "c(answer,nextNumberToCheck)" would not 
      # have had been defined.
      numberPrimesFoundSoFar <- numberPrimesFoundSoFar + 1  # If the number was a prime then add one to the
      # variable "numberPrimesFoundSoFar"
    }  # end of code for the "if"
    nextNumberToCheck <- nextNumberToCheck + 1  # Whether or not the last number was a prime, move on to
    # check the next number by adding one to "nextNumberToCheck"
  } # end of code for the "while"
  
  return (answer)
}
```
:::

---

## Matrices


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 70.   TPICS: user defined functions, matrices
#   For the following question refer to the following definitions:
#   A "square matrix" is a matrix that has the same number of rows and columns.
#
#   A "diagonal matrix" is a square matrix that has 1's (or another value) on the
#     "main diagonal" that runs from the upper left hand corner to the lower right
#     hand corner and zeros everywhere else. Examples:
#                               1  0  0     3 0 0 0
#                               0  1  0     0 3 0 0 
#                               0  0  1     0 0 3 0
#                                           0 0 0 3
# An "upper triangular" matrix is a square matrix that has non-zero values on
#   the main diagonal and in the upper right hand "triangle" of the matrix and
#   zeros everywhere else. Examples:
#                               1  1  1     3 3 3 3
#                               0  1  1     0 3 3 3 
#                               0  0  1     0 0 3 3
#                                           0 0 0 3
# A "lower triangular" matrix is a square matrix that has non-zero values on
#   the main diagonal and in the upper right hand "triangle" of the matrix and
#   zeros everywhere else. Examples:
#                               1  0  0     3 0 3 0
#                               1  1  0     3 3 0 0 
#                               1  1  1     3 3 3 0
#                                           3 3 3 3
# ARGUMENTS/PARAMETERS:
#
#     size - a single number that represents the number of rows (and hence
#            the number of columns) of a square matrix
#
#     shape - expected to be either "diagonal", "upperTriangular" or "lowerTriangular"
#
#     fillValue - a value that is used to fill the nonzero entries in the resulting matrix
#
# Returns a square matrix of the specified "size" and "shape". The nonzero entries in the
# matrix should be filled with the value of "fillValue"
#
# HINTS: See the R documentation for the row() and col() functions. Try a 
#        few examples of using the row() and col() functions. Then try to
#        use those functions to implement the solution to this question. 
#
# FUNCTION SIGNATURE: 
#     makeSquareMatrix <- function (size = 3 , shape = "diagonal" , fillValue = 1)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
makeSquareMatrix <- function (size = 3 , shape = "diagonal" , fillValue = 1) {
  mat <- matrix ( rep(0,size^2), nrow=size, ncol=size)
  
  if (shape == "upperTriangular") {
    
    mat[ row(mat) <= col(mat)] <- fillValue
    
  } else if (shape == "lowerTriangular") {
    
    mat[ row(mat) >= col(mat)] <- fillValue
    
  } else {
    mat [ row(mat) == col(mat) ] <- fillValue
  }
  
  mat
}

# Examples of using the function:
makeSquareMatrix(4,"upperTriangular", 99)
makeSquareMatrix(3,"lowerTriangular", 2)
makeSquareMatrix(3,"diagonal", 1)
makeSquareMatrix(3,"bogusValueForShape", 1)  # based on the code above this will create a "diagonal" matrix
```
:::

---


### QUESTION 71a.   TOPICS: user defined functions, vectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 71a.   TOPICS: user defined functions, vectors
#
# This function returns the "middle" values from a vector in the same order
# that they appear in the vector (see the examples below).
#
# ARGUMENTS/PARAMETERS:
#
#    data - a vector
#    n   - the number of middle values to return. This defaults 
#          to the entire vector.
#
# Examples: 
#    > middle.values( c(40,30,20,10), 2)  # return the middle 2 values
#    30  20 
#
#    > middle.values( c(40,30,20,10))   # n wasn't specified so default is entire vector
#    40  30  20  10
#
#    > # If there aren't exactly "n" "middle values" (i.e. if n is odd and length(data) is even
#    > # or n is even and length(data) is odd) then the function should return
#    > # one more value from the first half of "data" and one fewer value from
#    > # the second half of "data". For example in the following, 
#    > # there isn't just one middle value - there are two, so return the first "middle value".
#    >
#    > middle.values( c(40,30,20,10), 1)  
#    30
#
#    > # similarly:
#    > middle.values( c(40,30,20,10), 3)    # rerurn the 40, but not the 10
#    40  30  20
#
#    > # SOME OTHER EXAMPLES:
#    > middle.values( c(2,9,3,8,4,7,5), 4)
#    9  3  8  4
#    > middle.values( c(2,9,3,8,4,7,5), 2)
#    3  8 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
middle.values <- function ( data , n = length(data)) {
 
  middlePosition = length(data) %/% 2  
  
  startOfMiddleValues = middlePosition - n / 2 + 1
  
  endOfMiddleValues = startOfMiddleValues + n -1
  
  data[ startOfMiddleValues:endOfMiddleValues ]
  
}

middle.values( c(40,30,20,10), 2)   # 30  20 

middle.values( c(40,30,20,10))      # 40  30  20  10

middle.values( c(40,30,20,10), 1)   # 30

middle.values( c(40,30,20,10), 3)   # 40  30  20

middle.values( c(2,9,3,8,4,7,5), 4) # 9  3  8  4

middle.values( c(2,9,3,8,4,7,5), 2) # 3  8 
```
:::

---


### 71 b. TOPICS: user defined functions, vectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION
# 71 b. 
# TOPICS: user defined functions, vectors
#
# Modify the middle.values function from the previous question so that it takes 
# an additional argument named "percent". "percent" specifies the percent of
# values that should be returned as the middle values. The value of "percent" 
# should only be used if "n" is not specified. If both "n" and "percent" are
# specified then the value of "n" should be used. If neither "n" nor "percent"
# are specified then return all of the values.
#
# ARTUMENTS/PARAMETERS:
#    n       - the number of middle values to be returned
#    percent = a number between 0 and 1 (ie. 1 is 100 percent)
#
# EXAMPLES:
#   > # return the middle 40% of the values
#   > middle.values( c(100,90,80,70,60,50,40,30,20,10),  percent=0.4)
#   70 60 50 40
#
#   > # return the middle 6 values
#   > middle.values( c(100,90,80,70,60,50,40,30,20,10),  n=6)
#   80 70 60 50 40 30
#
#   > # both n and percent are specified, so use "n" and NOT "percent"
#   > middle.values( c(100,90,80,70,60,50,40,30,20,10),  n=6, percent=0.4)
#   80 70 60 50 40 30
#
#   > # neither n nor percent are specified, so return ALL of the values.
#   > middle.values( c(100,90,80,70,60,50,40,30,20,10),  n=6, percent=0.4)
#   100 90 80 70 60 50 40 30 20 10 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
middle.values <- function ( data , n = NA, percent = NA)
# TODO - finish this
```
:::

---

### QUESTION 71c.


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 71c.
#
# Modify the middle.value function from the previous question that that it takes 
# an additional argument named "sort".
#
# If the value of "sort" is specified as  "asc" then the middle values that
# are returned should be sorted in "ascending" order (e.g. 1,2,3,...)
# 
# If the value of "sort" is specified as  "desc" then the middle values that
# are returned should be sorted in "descending" order (e.g. 10,9,8,,...)
#
# If the value of "sort" is specified as  "none" (or as any other value) 
# then the middle values should be returned in the same order in which they 
# appear in the original vector.
# 
# EXAMPLES:
#     > middle.values(c(10,1,9,2,8,3), 4,  sort = "asc" )
#     1  2  8  9
#
#     > middle.values(c(10,1,9,2,8,3), 4,  sort = "desc" )
#     9  8  1  2
#
#     > middle.values(c(10,1,9,2,8,3), 4,  sort = "none" )
#     1  9  2  8
#
#     > middle.values(c(10,1,9,2,8,3), 4 )   # default is sort = "none"
#     1  9  2  8
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
middle.values <- function ( data , n = NA, percent = NA, sort="none")     
# TODO - finish this  
```
:::

---

### 71. d. TOPICS: user defined functions


```{r}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION
# 71. d.
# TOPICS: user defined functions
#
# Modify the middle.values function from the previous question so that the "data"
# argument may be a vector, a dataframe or a generic list that contains several vectors. 
#
# If data is a vector then this function should work exactly the same way as
# described in the previous question.
#
# If data is a dataframe then the function should return a dataframe that contains 
# the middle values (as specified) for each column of the dataframe. 
#
# If data is a generic list, thet contains several vecotrs, then the function should
# return a list that contains several vectors. Each vector in the returned list 
# should contain the middle values from the corresponding vector in the original
# list.
# 
# EXAMPLES:
#     You should be able to understand the question from the description above,
#     but I'll try to post some examples later.
#
# HINTS:  
#    - create two functions, the first function should do what the previous
#      question required. The 2nd function should be the function described here.
#      This 2nd function (i.e. the one that you need to write for THIS question) 
#      should "apply" the first function to the columns of a dataframe or the elements
#      of a generic list as necessary by using one of the "apply" functions.
#    - the is.data.frame() function can be used to determine if a variable contains a dataframe.
#    - mode() can be used to determine if a variable contains a list. Note that a "dataframe"
#      is also technically a "list".
#    - use if/else if/else    as necessary
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
middle.values <- function ( data , n = NA, percent = NA, sort="none")
# TODO - finish this
```
:::

---


### QUESTION 71e. TOPICS: user defined functions

```{r}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 71e.
# TOPICS: user defined functions
#
# Write a function that returns a vector of the means of different increasing
# amounts of "middle values" from a vector. If the number of values in the original
# vector are even, then the means returned should start with the means of the 2 
# middle values. If the number of values in the original vector are odd, then the
# means should start with the means of the 1 middle value. For example, 
# 
#    > # even number of values in vector
#    > get.middle.means( c(5,4,1,2,3,6) )   # means of  1,2    4,1,2,3  and  5,4,1,2,3,6
#    1.5  2.5  5.25    
#
#    > # odd number of values in vector
#    > get.middle.means( c(5,4,1,2,3,6,7) ) # means of  2   1,2,3   4,1,2,3,6  and  5,4,1,2,3,6,7
#    2   2   3.2   4    
#
# HINTS: 
#  - use the middle.values function that you created above to get the middle values
#  - use a loop   (either a for loop or a while loop)  to keep getting different 
#    numbers of middle.values and getting the mean of those middle values.
#  - you "answer" can start as an "empty vector" by using the code:   answer <- vector()
#    Then add a new number to the answer vector each time your loop finds a new mean.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
get.middle.means = function( vec )
# TODO - finish this  
```
:::

---

## Loops

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 72a.   TOPICS: loops
# Write a function, numPairs, that takes a single argument, nums that is
# expected to be a numeric vector.  The function should return the number of
# pairs of numbers in nums that add up to a multiple of 60.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
numPairs <- function(nums){
  answer <- list()
  for(pos1 in 1:(length(nums)-1)){
    n1 <- nums[pos1]
    
    for(pos2 in (pos1+1):length(nums)){
      n2 <- nums[pos2]
      if( (n1+n2) %% 60 == 0 ){
        answer[[length(answer)+1]] <- c(n1,n2)
      }
    }
  }
  length(answer)
}

# check to see it works
amounts <- c(70,5,50,20,40,10,100)
x <- numPairs(amounts)
x
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 72b. 
# Modify the function, numPairs, so that it takes a 2nd argument, sumAmount.
# The default value of sumAmount should be 60. The function should return the
#  number of pairs of numbers in nums that add up to a multiple of sumAmount.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
numPairs <- function(nums, sumAmount=60){
  answer <- list()
  for(pos1 in 1:(length(nums)-1)){
    n1 <- nums[pos1]
    
    for(pos2 in (pos1+1):length(nums)){
      n2 <- nums[pos2]
      if( (n1+n2) %% sumAmount == 0 ){
        answer[[length(answer)+1]] <- c(n1,n2)
      }
    }
  }
  length(answer)
}

# check to see it works
amounts <- c(70,5,50,20,40,10,100)
x <- numPairs(amounts)
x
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 72c. 
# Write a function, getPairs, that takes a single argument, nums that is
# expected to be a numeric vector. The function should return a list that
# contains the pairs of numbers in nums that add up to a multiple of 60.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
numPairs <- function(nums){
  answer <- list()
  for(pos1 in 1:(length(nums)-1)){
    n1 <- nums[pos1]
    
    for(pos2 in (pos1+1):length(nums)){
      n2 <- nums[pos2]
      if( (n1+n2) %% 60 == 0 ){
        answer[[length(answer)+1]] <- c(n1,n2)
      }
    }
  }
  answer
}

# check to see it works
amounts <- c(70,5,50,20,40,10,100)
x <- numPairs(amounts)
x
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 72d. 
# Write a function, getPairs, that takes a single argument, nums that is 
# expected to be a numeric vector. The function should return a dataframe that
# contains the pairs of numbers in nums that add up to a multiple of 60.
# Each pair should be in a new column on the dataframe. The names of the columns
# should be pair1, pair2, pair3, etc.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
numPairs <- function(nums){
  answer <- list()
  for(pos1 in 1:(length(nums)-1)){
    n1 <- nums[pos1]
    
    for(pos2 in (pos1+1):length(nums)){
      n2 <- nums[pos2]
      if( (n1+n2) %% 60 == 0 ){
        answer[[length(answer)+1]] <- c(n1,n2)
      }
    }
  }
  answer = data.frame(answer)
  colnames(answer) = paste0("pair", 1:ncol(answer))
  answer
}

# check to see it works
amounts <- c(70,5,50,20,40,10,100)
x <- numPairs(amounts)
x
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 72e.
#
# Write a function, getPairs, that takes a single argument, nums that is
# expected to be a numeric vector. The function should return a matrix that
# is 2 rows by n columns where n is the number of pairs of numbers from nums
# that add up to 60. Each column of the matrix should contain a pair of numbers.
# The names of the columns should be pair1, pair2, pair3, etc
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
numPairs <- function(nums){
  answer <- numeric(0)
  for(pos1 in 1:(length(nums)-1)){
    n1 <- nums[pos1]
    
    for(pos2 in (pos1+1):length(nums)){
      n2 <- nums[pos2]
      if( (n1+n2) %% 60 == 0 ){
        answer[length(answer)+c(1,2)] <- c(n1,n2)
      }
    }
  }
  answer = matrix(answer, nrow=2, ncol=length(answer)/2)
  colnames(answer) = paste0("pair", 1:ncol(answer))
  answer
}

# check to see it works
amounts <- c(70,5,50,20,40,10,100)
x <- numPairs(amounts)
x
```
:::

---



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 72f.
#
# Write a function, getPairs, that takes a single argument, nums that is
# expected to be a numeric vector. The function should return a matrix that
# is n rows by 2 columns where n is the number of pairs of numbers from nums
# that add up to 60. Each row of the matrix should contain a pair of numbers.
# The names of the rows should be pair1, pair2, pair3, etc
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
numPairs <- function(nums){
  answer <- numeric(0)
  for(pos1 in 1:(length(nums)-1)){
    n1 <- nums[pos1]
    
    for(pos2 in (pos1+1):length(nums)){
      n2 <- nums[pos2]
      if( (n1+n2) %% 60 == 0 ){
        answer[length(answer)+c(1,2)] <- c(n1,n2)
      }
    }
  }
  answer = matrix(answer, nrow=length(answer)/2, ncol=2, byrow = TRUE)
  rownames(answer) = paste0("pair", 1:nrow(answer))
  answer
}

# check to see it works
amounts <- c(70,5,50,20,40,10,100)
x <- numPairs(amounts)
x
```
:::

---

## Additional Questions - Assorted Topics

### QUESTION 73 TOPIC: vectors

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 73
# TOPIC: vectors
# 
# Show the last 5% of the values in a vector named v.
# For example 
# - if there are 100 values in v, show the last 5 values
# - if there are 200 values in v, show the last 10 values
# - etc.
# - if there are 43 values in v, 5% of 43 is 2.15 so round up (use ceiling function)
#   to show 3 last values
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
v = seq(10,1000,by=10)                                    # test data - 100 values
v [ (length(v) - ceiling(.05*length(v)) + 1):length(v) ]  # ANSWER

v = seq(10,2000,by=10)                                       # test data - 200 values
v [ (length(v) - ceiling(.05*length(v)) + 1):length(v) ]     # ANSWER (same as above)

v = seq(10,430,by=10)                                        # test data - 43 values
v [ (length(v) - ceiling(.05*length(v)) + 1):length(v) ]     # ANSWER (same as above)
```
:::

---


### QUESTION 74 TOPIC: dataframes

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 74
# TOPIC: dataframes
#
# Show the last 5% of the rows of the data in the dataframe named df.
# For example 
# - if there are 100 rows, show the last 5 rows
# - if there are 200 rows, show the last 10 rows
# - etc.
# - if there are 43 rows, 5% of 43 is 2.15 so round up (use ceiling function)
#   to show 3 last rows
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# TODO
```
:::

---

### QUESTION 75 TOPIC: user defined functions

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 75
# TOPIC: user defined functions (creating your own functions)
#
# Write a function with the "signature", evensBetween = function(x,y). 
# x and y are expected to be numbers (they may be even or odd).
# The function should return the even numbers between x and y (including x and y)
# You may assume that the value of x is less than the value of y.
#
# For example:
# > evensBetween(3,9)    
# [1] 4 6 8
#
# > evensBetween(2,10)
# [1] 2 4 6 8 10
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
evensBetween = function(x,y){
  numsBetweenXandY = x:y
  numsBetweenXandY[ numsBetweenXandY %% 2 == 0 ]
}

# test it:
evensBetween(3,9)    # 4 6 8
evensBetween(2,10)   # 2 4 6 8 10
```
:::

---

### QUESTION 76 TOPIC: user defined functions

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 76
# TOPIC: user defined functions
# SEE ALSO: question 11 for a similar question that doesn't require the creation of a function
#
# Create a function with the signature, swapFirstAndLast = function(vec,num).
# The vector vec is expected to contain at least 2*num values.
# The function should return a vector that contains the same values as vec except
# the first "num" values in vec should be last and the last "num" values in vec should be first
# (see the examples). 
#
# Examples:  
#       > swapFirstAndLast( seq(10,100,by=10), 3)
#       [1] 80 90 100 40 50 60 70 10 20 30
#
#       > swapFirstAndLast( seq(10,100,by=10), 2)
#       [1] 90 100 30 40 50 60 70 80 10 20
#
#       > swapFirstAndLast( c("abe","bob","carla","dana","ed","frank"), 1)
#       [1] "frank"  "bob"  "carla"  "dana"  "ed"  "abe"
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
swapFirstAndLast = function(vec , num) {
  firstPositions = 1:num
  lastPositions = (length(vec)-num+1):length(vec) 
  middlePositions = (num+1):(length(vec)-num)
  vec[ c( lastPositions , middlePositions, firstPositions) ]
}

# Test the answer
swapFirstAndLast( seq(10,100,by=10), 3)
swapFirstAndLast( seq(10,100,by=10), 2)
swapFirstAndLast( c("abe","bob","carla","dana","ed","frank"), 1)
```
:::

---


### QUESTION  77 TOPIC: user defined functions (creating your own functions)


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION  77
# TOPIC: user defined functions (creating your own functions)
# (see question 11 for a similar question that doesn't require the creation of a function)
#
# Define a function with the signature:  collapsevector = function( vec )
# vec is expected to be a vector with an even number of entries.
#
# The function should return a vector that contains 
# the sum of the first two items from vec
# followed by the sum of the next two items from vec
# followed by the sum of the next two items from vec
# etc.
# 
# Examples:
#    > collapseVector( c(10,20,30,40,50,60,70,80) )
#    [1]  30  70  110  150
# 
#    > collapseVector( c(3,4,10,2,100,3))
#    [1]  7  12  103
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
collapseVector = function( nums ){
  firstThirdFifthEtc <- nums[seq(1,length(nums),by=2)]   # ANSWER (MULTIPLE LINES)
  secondFourthSixthEtc <- nums[seq(2,length(nums),by=2)]     # ANSWER (MULTIPLE LINES)
  answer <- firstThirdFifthEtc + secondFourthSixthEtc  # ANSWER (MULTIPLE LINES)
  answer
}

# Test the answer
collapseVector( c(10,20,30,40,50,60,70,80) )
collapseVector( c(3,4,10,2,100,3))
```
:::

---


### QUESTION 78 TOPICS:  user defined functions (creating your own functions)

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 78
# TOPICS:  user defined functions (creating your own functions)
#
# Define a function with the signature:  get2ndLowest = function( vec )
# that returns the 2nd lowest value in vec.
#
# EXAMPLES:
# > get2ndLowest(c(423,6,234,3,7,1))
# [1] 3
#
# > get2ndLowest(c(3,9,10,-23,-59,200))
# [1] -23
#
# > get2ndLowest( c("grapes", "apple", "plum", "banana", "pear") )
# [1] "banana" 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
get2ndLowest = function( x ){
  xWithoutLowestNumber = x[ x != min(x) ]
  answer = min(xWithoutLowestNumber) 
  answer
}

# Test the answer
get2ndLowest(c(423,6,234,3,7,1))    # 3 
get2ndLowest(c(3,9,10,-23,-59,200))        # -23
get2ndLowest( c("grapes", "apple", "plum", "banana", "pear") )  # "banana"
```
:::

---


### QUESTION 79 TOPICS: vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 79
# TOPICS: vectors
#
# x and y are two vectors of the same length 
# combine x and y into a new vector named answer that contains the odd position values from x
# and the even position values of y.
# EXAMPLES
#    > x = c(10,20,30,40)
#    > y = c(100, 200, 300, 400)
#    > # YOUR CODE GOES HERE (it can be a few lines if you like)
#    > answer
#   [1] 10 200 30 400
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
x = c(10,20,30,40)

y = c(100, 200, 300, 400)

answer = x                                                     # ANSWER MULTIPLE LINES  - set answer to x

answer[ seq(2,length(y), by=2) ] = y [ seq(2,length(y), by=2)] # ANSWER MULTIPLE LINES - set even positions to the even positions from y

answer # show the result
```
:::

---


### QUESTION 80   TOPICS: lists, loops, user defined functions


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 80
# TOPICS: lists, loops, user defined functions
# SEE ALSO: question 72 for a similar question that doesn't use loops
# 
# Write a function with the following "signature":
#
# SIGNATURE:  getAllPairs = function (vec)
#
# ARGUMENTS/PARAMETERS 
#
#   vec - a vector that contains 2 or more values
#
# RETURNS: a list that contains a separate vector for each pair of values from vec. 
#          The values in the vectors should be in increasing order of value.
#
# EXAMPLE 1: 
#
#   > getAllPairs(c(200, 100, 300))
#
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# TODO
```
:::

---


### QUESTION 81   TOPICS: user defined functions, vectors


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# QUESTION 81   TOPICS: user defined functions, vectors
#
# Write a function with the following "signature":
#
# SIGNATURE:  getTopTwoValues = function (vec)
#
# ARGUMENTS/PARAMETERS 
#
#   vec - a vector that contains 2 or more values
#
# Returns a vector that contains the two largest
# values from the vector,vec . The values should 
# be returned IN INCREASING ORDER regardless
# of the order in the original vector. For example:
#     > getTopTwoValues(c(1,4,2,5,3))
#     4  5
#     > getTopTwoValues(c(1,5,2,4,3))
#     4  5
# If the largest value in vec appears more than once
# then the return value should include that value twice.
# For example:
#     > getTopTwoValues(c(1,4,2,4,3,4))
#     4  4
#
# Hints: You can write the function in many different ways. 
# Some functions that might help are: sort, head, tail 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

::: {.callout-note icon=false collapse="true"}
#### click here for answer {.unnumbered}

```{r}
# TODO
```
:::

